### 2020-09-08 annotated-bibliography.tex
Add CSCI Senior Sem templates and start bibliography

% This is a sample document using the University of Minnesota, Morris, Computer Science
% Senior Seminar modification of the ACM sig-alternate style to generate a simple annotated
% bibliography. The idea is that this document is fairly short, consisting of a brief description
% of your sources and how you intend to use them (or not). Most of the ``content'' of the
% generated document comes from the bibliography file, including the notes field which will
% provide the annotations.

% See https://github.com/UMM-CSci/Senior_seminar_templates for more info and to make
% suggestions and corrections.

\documentclass{sig-alternate}

\usepackage{url}

\begin{document}

% --- Author Metadata here ---
%%% REMEMBER TO CHANGE THE SEMESTER AND YEAR AS NECESSARY
\conferenceinfo{UMM CSci Senior Seminar Conference, October 2020}{Morris, MN}

\title{Recent advances in smartphone computational photography}

\numberofauthors{1}

\author{
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
\alignauthor
Paul Friederichsen\\
	\affaddr{Division of Science and Mathematics}\\
	\affaddr{University of Minnesota, Morris}\\
	\affaddr{Morris, Minnesota, USA 56267}\\
	\email{fried701@morris.umn.edu}
}

\maketitle

\begin{abstract}
My topic is recent advances in the computational photography used in smartphones. Specifically looking at the use of multi-frame algorithms to increase image resolution and quality. I'm mainly looking at a couple of papers by researchers at Google on algorithms used in the Super-Res Zoom and Night Sight features of the Google Pixel series of smartphones.
\end{abstract}

\section{Discussion of sources}

\subsection{Sources I expect to use (and how)}

My primary two sources are \cite{Wronski2019, Liba2019}.

% The following two commands are all you need to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% annotated_bibliography.bib is the name of the BibTex file containing 
% all the bibliography entries for this example. Note that you *don't* include the .bib ending
% in the \bibliography command.
\bibliography{paper}

% You must have a ".bib" file and remember to run:
%     pdflatex bibtex pdflatex pdflatex
% in order to see all the citation references correctly.

\end{document}

### 2020-09-08 paper.tex
Add CSCI Senior Sem templates and start bibliography

% This is a sample document using the University of Minnesota, Morris, Computer Science
% Senior Seminar modification of the ACM sig-alternate style. Much of this content is taken
% directly from the ACM sample document illustrating the use of the sig-alternate class. Certain
% parts that we never use have been removed to simplify the example, and a few additional
% components have been added.

% See https://github.com/UMM-CSci/Senior_seminar_templates for more info and to make
% suggestions and corrections.

\documentclass{sig-alternate}
\usepackage{color}
\usepackage[colorinlistoftodos]{todonotes}

%%%%% Uncomment the following line and comment out the previous one
%%%%% to remove all comments
%%%%% NOTE: comments still occupy a line even if invisible;
%%%%% Don't write them as a separate paragraph
%\newcommand{\mycomment}[1]{}

\begin{document}

% --- Author Metadata here ---
%%% REMEMBER TO CHANGE THE SEMESTER AND YEAR AS NEEDED
\conferenceinfo{UMM CSci Senior Seminar Conference, April 2018}{Morris, MN}

\title{UMM CSci Senior Seminar LaTeX template}

\numberofauthors{1}

\author{
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
\alignauthor
Chris S. Student\\
	\affaddr{Division of Science and Mathematics}\\
	\affaddr{University of Minnesota, Morris}\\
	\affaddr{Morris, Minnesota, USA 56267}\\
	\email{cssxxxx00000@morris.umn.edu}
}

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for the University of
Minnesota, Morris, Computer Science Senior Seminar proceedings.
It is based heavily on (and takes material directly from) a similar document 
illustrating the format of the ACM SIG Proceedings, which we have based our
proceedings format on.

The original ACM document tried to include
\begin{quote}
every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.
\end{quote}

We've removed many of the more esoteric tricks here because either
they'll never be used (e.g., multiple authors) or are used \emph{very}
rarely (e.g., appendices). Refer to the original ACM document for more
of those fancy examples.
% The current paper format *only* allows inline comments using the todo
% macro. That's kind of a bummer, and it would be neat if someone figured
% out how to change the acmconf style to allow this. I suspect it isn't *hard*
% but there are quite a few details that have to be sorted out in synchrony.
\todo[inline]{Needs more work}
\end{abstract}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
\label{sec:introduction}

The \textit{proceedings} are the records of a conference, and conference
editors like ACM seek to give their conference by-products a uniform,
high-quality appearance. We also would like our proceedings to look highly
professional, so we're borrowing heavily from the ACM formatting guidelines.
These include some
rigid requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (2.54cm [1"] top and
bottom and 1.9cm [.75"] left and right; specified column width
(8.45cm [3.33"]) and gutter size (.083cm [.33"]).

\todo[inline, color=green]{I really like this section.}

The good news is, with only a handful of manual
settings\footnote{One of these, the \texttt{\textbackslash alignauthor}, command, 
you have
already used; another, \texttt{\textbackslash balancecolumns} will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.} the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands. Section~\ref{sec:body} introduces the main
examples of formatting, and Section~\ref{sec:conclusions} wraps things up.

\section{The {\secit Body} of The Paper}
\label{sec:body}

Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{\textbackslash section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
\label{sec:typeChangesSpecialChars}

We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{\textbackslash textit}; emboldening with the
command \texttt{\textbackslash textbf}
and typewriter-style (for instance, for computer code) with
\texttt{\textbackslash texttt}.
As a rule you'd prefer \texttt{\textbackslash emph} (which stands for \emph{emphasize})
over something like \texttt{\textbackslash textit} since that gives the typesetting system
more flexibility in how it can emphasize that text.

You do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}~\cite{Lamport:LaTeX}.

\subsection{Math Equations}
\label{sec:mathEquations}

You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
\label{sec:inlineEquations}

A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{\textbackslash begin\ldots\textbackslash end}
construction or with the short form \texttt{\$\ldots \$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
\label{sec:displayEquations}

A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:

\begin{equation*}
\lim_{n\rightarrow \infty}x=0
\end{equation*}

Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:

\begin{displaymath}
	\sum_{i=0}^{\infty} x + 1
\end{displaymath}

and follow it with a numbered equation (Equation~\ref{eq:summation}):

\begin{equation}
	\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\label{eq:summation}
\end{equation}

just to demonstrate \LaTeX's able handling of numbering.
Note that if you use numbered equations, you can give them \texttt{label}s
and \texttt{\textbackslash ref} them, e.g., Equation~\ref{eq:summation}.

\subsection{Multi-line formulas}
\label{sec:multiLineFormulas}

The \texttt{align} and \texttt{align*} environments let you align sequences of
equations so, for examples, a sequences of equal signs line up nicely.

% The & creates a "virtual tab stop" and corresponding &'s on each line
% will be aligned when the layout is done.
\begin{align*}
n_1 &= \sum_{i = 1}^k a_i \\
n_{x-y} &= \prod_{i = 1}^k b_i
\end{align*}

\begin{align*}
 f(x) &= (x+a)(x+b) \\
 &= x^2 + (a+b)x + ab
\end{align*}

\subsection{Figures}
\label{sec:figures}

Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of
a \textbf{.pdf} file to be displayable with \LaTeX, such as
Figure~\ref{fig:singleColumnFigure}.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\psfig{file=sample_graph.pdf,width =3in}
\caption{A sample graph just spanning one column.}
\label{fig:singleColumnFigure}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns, like Figure~\ref{fig:twoColumnFigure}.  
To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\texttt{figure*} to enclose the figure and its caption.
And don't forget to end the environment with
\texttt{figure*}, not \texttt{figure}!

\begin{figure*}
\centering
\psfig{file=sample_graph.pdf,width =5.3in}
\caption{A sample graph that needs to span two columns of text.}
\label{fig:twoColumnFigure}
\end{figure*}

It's easiest and you tend to get the best quality if your figure uses vector graphics
in PDF format. You can include other formats such as PNG, but they will usually
not look nearly as professional, especially when printed on high resolution printers.
\emph{Be very wary of screen captures from other papers. They tend to look pixelated
and amateurish even at high resolutions.}

\subsection{Tables}
\label{sec:tables}

Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:frequencyOfSpecialChars} is included in the input file; 
compare the placement of the table here with the table in the
PDF output running \LaTeX\ on this document.

\begin{table}[t]
\centering
\caption{Frequency of Special Characters}
\label{tab:frequencyOfSpecialChars}
\begin{tabular}{c|c|l}
Non-English or Math & Frequency & Comments\\ \hline
\O & 1 in 1,000 & For Swedish names\\
$\pi$ & 1 in 5 & Common in math\\
\$ & 4 in 5 & Used in business\\
$\Psi^2_1$ & 1 in 40,000 & Unexplained usage\\
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption, as demonstrated in Table~\ref{tab:typicalCommands} below.  
As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table~\ref{tab:typicalCommands} is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}[t]
\centering
\caption{Some Typical Commands}
\label{tab:typicalCommands}
\begin{tabular}{ccl}
Command & A Number & Comments \\ \hline
\texttt{\textbackslash alignauthor}     & 100 & Author alignment\\
\texttt{\textbackslash numberofauthors} & 200 & Author enumeration\\
\texttt{\textbackslash table}           & 300 & For tables\\
\texttt{\textbackslash table*}          & 400 & For wider tables\\ 
\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Citations}
\label{sec:citations}

Citations to articles~\cite{Aaronson:2005,Garey:1979,Brun:2008} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file~\cite{OM:2008}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

It is recommended that you precede \texttt{\textbackslash cite} (and 
\texttt{\textbackslash ref}) commands with a tilde
character instead of a space, e.g., \texttt{some text\textasciitilde\textbackslash cite}. The tilde gives you a non-breaking space which ensures that your citation won't get
stranded by itself on the beginning of a line.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}.

This article shows only the plainest form
of the citation command, using \texttt{\textbackslash cite},
which is all that is needed for our senior seminar.
You shouldn't use any other forms here.

\subsection{Theorem-like Constructs}
\label{sec:theoremLikeConstructs}

Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{\textbackslash newtheorem} and the
other by the command \texttt{\textbackslash newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

Theorem~\ref{thm:integration} below uses the \textbf{theorem} environment, created by
the \texttt{\textbackslash newtheorem} command:

% You would usually put a \newtheorem command up at the top 
% of your LaTeX document after the \usepackage commands. It's
% just here in this example so it's with the text that describes it.
\newtheorem{theorem}{Theorem}

\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\label{thm:integration}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{\textbackslash newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{\textbackslash newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{align*}
l &= \lim_{x\rightarrow c} f(x) \\
  &= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ] \\
  &= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)}  \\
  &= 0\cdot L  \\
  &= 0,
\end{align*}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition~\cite{salas:calculus} shown above,
use the \texttt{\textbackslash newtheorem} or the
\texttt{\textbackslash newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
\label{sec:caveatForExperts}

Because you have just been given permission to
use the \texttt{\textbackslash newdef} command to create a
new form, you might think you can
use \TeX's \texttt{\textbackslash def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{\textbackslash def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
\label{sec:conclusions}

This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.

\section*{Acknowledgments}
\label{sec:acknowledgments}

This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.

You want to use the \texttt{\textbackslash section*} version of the \texttt{section}
command, as an acknowledgments section typically does \emph{not} get
a number.

It is common (but by no means necessary) for students to thank
their advisor, and possibly other faculty, friends, and family who provided
useful feedback on the paper as it was being written.

In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% sample_paper.bib is the name of the BibTex file containing the
% bibliography entries. Note that you *don't* include the .bib ending here.
\bibliography{paper}  
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

\end{document}

### 2020-09-10 annotated-bibliography.tex
Updates from Overleaf

@@
 
-\subsection{Sources I expect to use (and how)}
+\subsection{Primary Sources}
 
 My primary two sources are \cite{Wronski2019, Liba2019}.
+\begin{itemize}
+\item \cite{Wronski2019} describes an algorithm to use natural hand tremor to create higher resolution images from multi-frame bursts.
+It has the effect of improving image resolution and noise performance.
 
+\item \cite{Liba2019} describes a mobile camera solution leveraging existing burst imaging techniques and improving low-light performance.
+It builds upon work done in \cite{Hasinoff2016} and uses the burst merging algorithm from \cite{Wronski2019}.
+\end{itemize}
+Both of these primary sources are long and very detailed. They are written by researchers at Google.
+
+\subsection{Background Sources}
+\begin{itemize}
+\item \cite{Hasinoff2016} is cited in both of the previous mentioned articles and describes an earlier multi-frame burst merging algorithm.
+It is used as a point of comparison by both primary sources and will be necessary background information.
+It is also from Google Research.
+I think this could end up being more of a primary source.
+
+\end{itemize}
 % The following two commands are all you need to

### 2020-09-10 annotated-bibliography.tex
Updates from Overleaf

@@
 \begin{itemize}
-\item \cite{Wronski2019} describes an algorithm to use natural hand tremor to create higher resolution images from multi-frame bursts.
+\item Wronski 2019 \cite{Wronski2019} describes an algorithm to use natural hand tremor to create higher resolution images from multi-frame bursts.
 It has the effect of improving image resolution and noise performance.
 
-\item \cite{Liba2019} describes a mobile camera solution leveraging existing burst imaging techniques and improving low-light performance.
+\item Liba 2019 \cite{Liba2019} describes a mobile camera solution leveraging existing burst imaging techniques and improving low-light performance.
 It builds upon work done in \cite{Hasinoff2016} and uses the burst merging algorithm from \cite{Wronski2019}.
@@
 \begin{itemize}
-\item \cite{Hasinoff2016} is cited in both of the previous mentioned articles and describes an earlier multi-frame burst merging algorithm.
+\item Hasinof 2016\cite{Hasinoff2016} is cited in both of the previous mentioned articles and describes an earlier multi-frame burst merging algorithm.
 It is used as a point of comparison by both primary sources and will be necessary background information.
@@
 I think this could end up being more of a primary source.
+\item Barron 2017 \cite{Barron2017} is on the auto-white-balance algorithm used in \cite{Liba2019} and will be used for background information on that.
+\item Wikipedia articles \cite{wiki:BayerFilter, wiki:Demosaicing, wiki:Aliasing} provide background information on Bayer filters, demosaicing, and aliasing that are necessary to explain how the algorithm in \cite{Wronski2019} improves on existing techniques.
+
+\item \cite{blog:Wronski2018, blog:Levoy2018} are the blog posts accompanying \cite{Wronski2019, blog:Levoy2018} respectively. They provide additional information and examples of the algorithms.
 

### 2020-09-17 annotated-bibliography.tex
Updates from Overleaf

@@
 \begin{itemize}
-\item Hasinof 2016\cite{Hasinoff2016} is cited in both of the previous mentioned articles and describes an earlier multi-frame burst merging algorithm.
+\item Hasinof 2016 \cite{Hasinoff2016} is cited in both of the previous mentioned articles and describes an earlier multi-frame burst merging algorithm.
 It is used as a point of comparison by both primary sources and will be necessary background information.
@@
 
-\item \cite{blog:Wronski2018, blog:Levoy2018} are the blog posts accompanying \cite{Wronski2019, blog:Levoy2018} respectively. They provide additional information and examples of the algorithms.
+\item \cite{blog:Wronski2018, blog:Levoy2018} are the blog posts accompanying \cite{Wronski2019, Liba2019} respectively. They provide additional information and examples of the algorithms.
 

### 2020-09-17 outline.tex
Updates from Overleaf

\documentclass{sig-alternate}

\usepackage{color}

\setlength{\marginparwidth}{2cm}
\usepackage[colorinlistoftodos]{todonotes}

\usepackage{url}

\begin{document}

% --- Author Metadata here ---
%%% REMEMBER TO CHANGE THE SEMESTER AND YEAR AS NECESSARY
\conferenceinfo{UMM CSci Senior Seminar Conference, October 2020}{Morris, MN}

\title{Recent advances in smartphone computational photography}

\numberofauthors{1}

\author{
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
\alignauthor
Paul Friederichsen\\
	\affaddr{Division of Science and Mathematics}\\
	\affaddr{University of Minnesota, Morris}\\
	\affaddr{Morris, Minnesota, USA 56267}\\
	\email{fried701@morris.umn.edu}
}

\maketitle

\begin{abstract}
My topic is recent advances in the computational photography used in smartphones. Specifically looking at the use of multi-frame algorithms to increase image resolution and quality. I'm mainly looking at a couple of papers by researchers at Google on algorithms used in the Super-Res Zoom and Night Sight features of the Google Pixel series of smartphones.
\todo[inline]{Not really an abstract at this point. Subject to change.}
\end{abstract}

\section*{\uppercase{Key points}}

This paper addresses how we can improve smartphone camera photography. Phone cameras present many challenges, most of which come from the need for them to be physically small. Their small size puts a fundamental limit on their ability to resolve detail and especially affects low-light photography. This paper will look at two approaches to improve smartphone photography through software techniques.

The first of two primary sources is titled ``Handheld Multi-Frame Super-Resolution" \cite{Wronski2019}. It describes a new method for merging bursts of raw images using the natural motion of the hand holding the device (or by moving the sensor in such a way to emulate that). One of the major contributions from this approach is the use of natural hand motion to obtain increased detail. It builds on work done in one of my background sources, ``Burst photography for high dynamic range and low-light imaging on mobile cameras" \cite{Hasinoff2016}.

The second primary source is ``Handheld Mobile Photography in Very Low Light" \cite{Liba2019}. This paper describes a new photography system (or pipeline) that improves the ability to capture photos in low light. This is done mainly through improvements in motion metering, motion-aware burst merging, and low-light optimized auto white balance. This system also builds on work done in \cite{Hasinoff2016}.

Both primary sources provide good testing and comparisons in their results sections. Both systems are in commercial use in Google Pixel phones.

I will probably need to assume the audience has a basic knowledge of how digital images are formed (pixels, RGB) but I will need to introduce more specific things like Bayer filters and demosaicing as well as how burst capturing works in modern smartphones.

\section{Introduction}

Introduce the problem-space and the basics of the two approaches that will be discussed.

% ----------------------------------------
\section{Background}

Include some basic background information on some terms that will be important.

\subsection{Burst photography}

This is probably the most important part of the background information. I will introduce the general idea of burst photography and the prior work both of my primary sources build on from \cite{Hasinoff2016}.

\subsubsection{Burst processing pipeline}

\subsection{Bayer filters}

Background information on Bayer filters. Some of this will come from \cite{wiki:BayerFilter}.

\subsection{Demosaicing}

Background information on demosaicing. Some of this will come from \cite{wiki:Demosaicing}.

\subsection{Aliasing}

Background information on aliasing, specifically in regard to signal processing. Some of this will come from \cite{wiki:Aliasing}.

% ----------------------------------------
\section{Handheld super-resolution}

This section is on the first primary source, \cite{Wronski2019}.

\subsection{Algorithm overview}

This will describe the general structure of the approach including frame acquisition, registration and alignment, and merging. The merging algorithm itself is explained in more detail later.

\subsection{Hand movement based super-resolution}

This section goes over the research on hand motion and why it is able to be used for super-resolution.

\subsection{Proposed super-resolution approach}

This will be the main part for this section and will describe the specifics of the algorithm proposed in \cite{Wronski2019}.

\subsubsection{Kernel reconstruction}

\subsubsection{Motion Robustness}

\subsection{Results}

A discussion of the results and comparisons discussed in the paper.

\subsubsection{Limitations}

% ----------------------------------------
\section{Handheld low light photography}

This section is on the second primary source, \cite{Liba2019}.

\subsection{Motion Metering}

Selecting exposure time based on a prediction of future movement.

\subsection{Motion-adaptive burst merging}

Builds on the Fourier domain temporal merging technique in \cite{Hasinoff2016} and improves the merging algorithm for low-light.

\subsection{Auto white balance in low-light}

Uses a new dataset to create a Fast Fourier Color Constancy algorithm for better white balance in low-light images.

\subsection{Tone mapping}

Looks at how to do tonemapping for low-light images so that they appear natural to humans.

\subsection{Results}

A discussion of the results and comparisons discussed in the paper.

\subsubsection{Limitations}

% ----------------------------------------
\section{Conclusions}

Some conclusion on how both systems are currently in use and future work that can be done with them.

\section{Acknowledgements}

% Bibliography
% ----------------------------------------

% Primary Sources
\nocite{Wronski2019, Liba2019}

% Background: supporting papers for primary sources
\nocite{Hasinoff2016, Barron2017}

% Background: Wikipedia articles
\nocite{wiki:BayerFilter, wiki:Demosaicing, wiki:Aliasing}

% Background: Blogs accompanying primary sources
\nocite{blog:Wronski2018, blog:Levoy2018}

% The following two commands are all you need to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% annotated_bibliography.bib is the name of the BibTex file containing 
% all the bibliography entries for this example. Note that you *don't* include the .bib ending
% in the \bibliography command.
\bibliography{paper}

% You must have a ".bib" file and remember to run:
%     pdflatex bibtex pdflatex pdflatex
% in order to see all the citation references correctly.

\end{document}

### 2020-09-24 paper.tex
Section Draft

@@
 \usepackage{color}
+
+\setlength{\marginparwidth}{2cm}
 \usepackage[colorinlistoftodos]{todonotes}
+
+\usepackage{url}
 
@@
 %%% REMEMBER TO CHANGE THE SEMESTER AND YEAR AS NEEDED
-\conferenceinfo{UMM CSci Senior Seminar Conference, April 2018}{Morris, MN}
+\conferenceinfo{UMM CSci Senior Seminar Conference, October 2020}{Morris, MN}
 
-\title{UMM CSci Senior Seminar LaTeX template}
+\title{Recent advances in smartphone computational photography}
 
@@
 \alignauthor
-Chris S. Student\\
+Paul Friederichsen\\
 	\affaddr{Division of Science and Mathematics}\\
@@
 	\affaddr{Morris, Minnesota, USA 56267}\\
-	\email{cssxxxx00000@morris.umn.edu}
+	\email{fried701@morris.umn.edu}
 }
@@
 \maketitle
+
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms,
-somewhat loosely, to the formatting guidelines for the University of
-Minnesota, Morris, Computer Science Senior Seminar proceedings.
-It is based heavily on (and takes material directly from) a similar document 
-illustrating the format of the ACM SIG Proceedings, which we have based our
-proceedings format on.
+(This is currently just a topic statement more than an abstract)
 
-The original ACM document tried to include
-\begin{quote}
-every imaginable sort
-of ``bells and whistles", such as a subtitle, footnotes on
-title, subtitle and authors, as well as in the text, and
-every optional component (e.g. Acknowledgments, Additional
-Authors, Appendices), not to mention examples of
-equations, theorems, tables and figures.
-\end{quote}
-
-We've removed many of the more esoteric tricks here because either
-they'll never be used (e.g., multiple authors) or are used \emph{very}
-rarely (e.g., appendices). Refer to the original ACM document for more
-of those fancy examples.
-% The current paper format *only* allows inline comments using the todo
-% macro. That's kind of a bummer, and it would be neat if someone figured
-% out how to change the acmconf style to allow this. I suspect it isn't *hard*
-% but there are quite a few details that have to be sorted out in synchrony.
-\todo[inline]{Needs more work}
+My topic is recent advances in the computational photography used in smartphones.
+Specifically looking at the use of multi-frame algorithms to increase image resolution and quality.
+I'm mainly looking at a couple of papers by researchers at Google on algorithms used in the Super-Res Zoom and Night Sight features of the Google Pixel series of smartphones.
 \end{abstract}
 
-\keywords{ACM proceedings, \LaTeX, text tagging}
+\keywords{computational photography, super-resolution, image processing, low-light imaging}
 
 \section{Introduction}
-\label{sec:introduction}
 
-The \textit{proceedings} are the records of a conference, and conference
-editors like ACM seek to give their conference by-products a uniform,
-high-quality appearance. We also would like our proceedings to look highly
-professional, so we're borrowing heavily from the ACM formatting guidelines.
-These include some
-rigid requirements for the format of the proceedings documents: there
-is a specified format (balanced  double columns), a specified
-set of fonts (Arial or Helvetica and Times Roman) in
-certain specified sizes (for instance, 9 point for body copy),
-a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
-the page, specified size of margins (2.54cm [1"] top and
-bottom and 1.9cm [.75"] left and right; specified column width
-(8.45cm [3.33"]) and gutter size (.083cm [.33"]).
+Smartphone cameras present many challenges, most of which come from the need for them to be physically small.
+Their small size puts a fundamental limit on their ability to resolve detail and especially affects low-light photography.
+This paper will look at two approaches to improve smartphone photography through software techniques.
 
-\todo[inline, color=green]{I really like this section.}
+% ----------------------------------------
+\section{Background}
 
-The good news is, with only a handful of manual
-settings\footnote{One of these, the \texttt{\textbackslash alignauthor}, command, 
-you have
-already used; another, \texttt{\textbackslash balancecolumns} will
-be used in your very last run of \LaTeX\ to ensure
-balanced column heights on the last page.} the \LaTeX\ document
-class file handles all of this for you.
+\subsection{Burst photography}
 
-The remainder of this document is concerned with showing, in
-the context of an ``actual'' document, the \LaTeX\ commands
-specifically available for denoting the structure of a
-proceedings paper, rather than with giving rigorous descriptions
-or explanations of such commands. Section~\ref{sec:body} introduces the main
-examples of formatting, and Section~\ref{sec:conclusions} wraps things up.
+Most smartphones use a burst processing pipeline for their cameras.
+Generally, burst processing involves taking a series of raw exposures and merging them together to form the final image.
+Most smartphones operate in a \emph{Zero-Shutter Lag} mode by default.
+In this mode raw frames are continuously captured to a ring buffer while the camera app is open.
+When the user presses the shutter button, a number of the most recent frames are sent to the camera processing pipeline.
 
-\section{The {\secit Body} of The Paper}
-\label{sec:body}
+Both of the approaches in this paper build on the end-to-end burst processing pipeline from Hasinoff et al.
+\cite{Hasinoff2016} which used bursts of constant low-exposure frames to increase dynamic range and signal-to-noise ratio.
+% TODO: more on Hasinoff
 
-Typically, the body of a paper is organized
-into a hierarchical structure, with numbered or unnumbered
-headings for sections, subsections, sub-subsections, and even
-smaller sections.  The command \texttt{\textbackslash section} that
-precedes this paragraph is part of such a
-hierarchy.\footnote{This is the second footnote.  It
-starts a series of three footnotes that add nothing
-informational, but just give an idea of how footnotes work
-and look. It is a wordy one, just so you see
-how a longish one plays out.} \LaTeX\ handles the numbering
-and placement of these headings for you, when you use
-the appropriate heading commands around the titles
-of the headings.  If you want a sub-subsection or
-smaller part to be unnumbered in your output, simply append an
-asterisk to the command name.  Examples of both
-numbered and unnumbered headings will appear throughout the
-balance of this sample document.
+\newpage
 
-Because the entire article is contained in
-the \textbf{document} environment, you can indicate the
-start of a new paragraph with a blank line in your
-input file; that is why this sentence forms a separate paragraph.
+\subsection{Bayer filter}
 
-\subsection{Type Changes and {\subsecit Special} Characters}
-\label{sec:typeChangesSpecialChars}
-
-We have already seen several typeface changes in this sample.  You
-can indicate italicized words or phrases in your text with
-the command \texttt{\textbackslash textit}; emboldening with the
-command \texttt{\textbackslash textbf}
-and typewriter-style (for instance, for computer code) with
-\texttt{\textbackslash texttt}.
-As a rule you'd prefer \texttt{\textbackslash emph} (which stands for \emph{emphasize})
-over something like \texttt{\textbackslash textit} since that gives the typesetting system
-more flexibility in how it can emphasize that text.
-
-You do not
-have to indicate typestyle changes when such changes are
-part of the \textit{structural} elements of your
-article; for instance, the heading of this subsection will
-be in a sans serif\footnote{A third footnote, here.
-Let's make this a rather short one to
-see how it looks.} typeface, but that is handled by the
-document class file. Take care with the use
-of\footnote{A fourth, and last, footnote.}
-the curly braces in typeface changes; they mark
-the beginning and end of
-the text that is to be in the different typeface.
-
-You can use whatever symbols, accented characters, or
-non-English characters you need anywhere in your document;
-you can find a complete list of what is
-available in the \textit{\LaTeX\
-User's Guide}~\cite{Lamport:LaTeX}.
-
-\subsection{Math Equations}
-\label{sec:mathEquations}
-
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of
-the three are discussed in the next sections.
-
-\subsubsection{Inline (In-text) Equations}
-\label{sec:inlineEquations}
-
-A formula that appears in the running text is called an
-inline or in-text formula.  It is produced by the
-\textbf{math} environment, which can be
-invoked with the usual \texttt{\textbackslash begin\ldots\textbackslash end}
-construction or with the short form \texttt{\$\ldots \$}. You
-can use any of the symbols and structures,
-from $\alpha$ to $\omega$, available in
-\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a
-few examples of in-text equations in context. Notice how
-this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsubsection{Display Equations}
-\label{sec:displayEquations}
-
-A numbered display equation -- one set off by vertical space
-from the text and centered horizontally -- is produced
-by the \textbf{equation} environment. An unnumbered display
-equation is produced by the \textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols
-and structures available in \LaTeX; this section will just
-give a couple of examples of display equations in context.
-First, consider the equation, shown as an inline equation above:
-
-\begin{equation*}
-\lim_{n\rightarrow \infty}x=0
-\end{equation*}
-
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-
-\begin{displaymath}
-	\sum_{i=0}^{\infty} x + 1
-\end{displaymath}
-
-and follow it with a numbered equation (Equation~\ref{eq:summation}):
-
-\begin{equation}
-	\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
-\label{eq:summation}
-\end{equation}
-
-just to demonstrate \LaTeX's able handling of numbering.
-Note that if you use numbered equations, you can give them \texttt{label}s
-and \texttt{\textbackslash ref} them, e.g., Equation~\ref{eq:summation}.
-
-\subsection{Multi-line formulas}
-\label{sec:multiLineFormulas}
-
-The \texttt{align} and \texttt{align*} environments let you align sequences of
-equations so, for examples, a sequences of equal signs line up nicely.
-
-% The & creates a "virtual tab stop" and corresponding &'s on each line
-% will be aligned when the layout is done.
-\begin{align*}
-n_1 &= \sum_{i = 1}^k a_i \\
-n_{x-y} &= \prod_{i = 1}^k b_i
-\end{align*}
-
-\begin{align*}
- f(x) &= (x+a)(x+b) \\
- &= x^2 + (a+b)x + ab
-\end{align*}
-
-\subsection{Figures}
-\label{sec:figures}
-
-Like tables, figures cannot be split across pages; the
-best placement for them
-is typically the top or the bottom of the page nearest
-their initial cite.  To ensure this proper ``floating'' placement
-of figures, use the environment
-\textbf{figure} to enclose the figure and its caption.
-
-This sample document contains examples of
-a \textbf{.pdf} file to be displayable with \LaTeX, such as
-Figure~\ref{fig:singleColumnFigure}.  More
-details on each of these is found in the \textit{Author's Guide}.
+The majority of digital image sensors in digital cameras and phones use a Bayer filter mosaic pattern to arrange RGB color filters on the sensor.
+The pattern consists of 50\% green, 25\% red, and 25\% blue pixels \cite{wiki:BayerFilter}.
+This ratio emulates the color sensitivity of the human eye.
+Due to this pattern on the sensor, the raw output of a digital camera also has each pixel filtered to only red, green, or blue and a demosaicing algorithm must be used to interpolate the other values for each pixel \cite{wiki:BayerFilter}.
 
@@
 \centering
-\psfig{file=sample_graph.pdf,width =3in}
-\caption{A sample graph just spanning one column.}
-\label{fig:singleColumnFigure}
+\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
+\caption{A Bayer pattern on a sensor in isometric perspective \cite{wiki:BayerFilter}.}
+\label{fig:BayerPattern}
 \end{figure}
 
-As was the case with tables, you may want a figure
-that spans two columns, like Figure~\ref{fig:twoColumnFigure}.  
-To do this, and still to
-ensure proper ``floating'' placement of tables, use the environment
-\texttt{figure*} to enclose the figure and its caption.
-And don't forget to end the environment with
-\texttt{figure*}, not \texttt{figure}!
+\subsection{Demosaicing}
 
-\begin{figure*}
+Background information on demosaicing. Some of this will come from \cite{wiki:Demosaicing}.
+
+The demosaicing process may introduce various artifacts in the final image. These typically include false color artifacts like zippering and Moiré patterns \cite{Wronski2019}.
+
+\subsection{Aliasing}
+
+Background information on aliasing, specifically in regard to signal processing. Some of this will come from \cite{wiki:Aliasing}.
+
+\subsection{Super-resolution}
+
+% ----------------------------------------
+\section{Handheld super-resolution}
+
+\begin{figure*}[t!]
 \centering
-\psfig{file=sample_graph.pdf,width =5.3in}
-\caption{A sample graph that needs to span two columns of text.}
-\label{fig:twoColumnFigure}
+% TODO: Size correctly for width
+\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
+\caption{An overview of the approach used by Wronski et al. \cite{Wronski2019}}
+\label{fig:Wronski2019Fig2}
 \end{figure*}
 
-It's easiest and you tend to get the best quality if your figure uses vector graphics
-in PDF format. You can include other formats such as PNG, but they will usually
-not look nearly as professional, especially when printed on high resolution printers.
-\emph{Be very wary of screen captures from other papers. They tend to look pixelated
-and amateurish even at high resolutions.}
+Wronski et al.~\cite{Wronski2019} introduce an algorithm that uses multiple shifted frames to produce higher resolution images from bursts of underexposed raw frames as part of the smartphone's imaging pipeline.
+The algorithm is able to directly use Bayer raw frames and removes the need for an explicit demosaicing step in the pipeline.
+It uses natural hand motion and is efficient enough to work in the background on smartphones.
 
-\subsection{Tables}
-\label{sec:tables}
+\subsection{Algorithm overview}
 
-Because tables cannot be split across pages, the best
-placement for them is typically the top of the page
-nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and
-the table caption.  The contents of the table itself must go
-in the \textbf{tabular} environment, to
-be aligned properly in rows and columns, with the desired
-horizontal and vertical rules.  Again, detailed instructions
-on \textbf{tabular} material
-is found in the \textit{\LaTeX\ User's Guide}.
+Wronski et al.'s approach is a process that starts with the the acquisition of a burst from the continous ring buffer of raw frames in the phone's camera application.
+Next, a single frame is chosen and the rest are aligned to it using a refined version of the algorithm by Hasinoff et al.~\cite{Hasinoff2016}.
+% TODO: elaborate on the alignment algorithm
+Each frame's local contributions are estimated through kernel regression (Section~\ref{sec:kernelReconstruction}) and accumulated across an whole burst for each of the three color planes.
+Next, in parallel, the kernel shapes are adjusted based on estimated local gradients and the sample contributions are adjusted weighted based on a robustness model (Section~\ref{sec:robustnessModel}).
+The final RGB image is obtained by normalizing the accumulated contributions for each of the three color planes and merging them together.
 
-Immediately following this sentence is the point at which
-Table~\ref{tab:frequencyOfSpecialChars} is included in the input file; 
-compare the placement of the table here with the table in the
-PDF output running \LaTeX\ on this document.
+\subsection{Hand movement based super-resolution}
 
-\begin{table}[t]
-\centering
-\caption{Frequency of Special Characters}
-\label{tab:frequencyOfSpecialChars}
-\begin{tabular}{c|c|l}
-Non-English or Math & Frequency & Comments\\ \hline
-\O & 1 in 1,000 & For Swedish names\\
-$\pi$ & 1 in 5 & Common in math\\
-\$ & 4 in 5 & Used in business\\
-$\Psi^2_1$ & 1 in 40,000 & Unexplained usage\\
-\end{tabular}
-\end{table}
+One of the important conditions for multi-frame super-resolution is that the input contains multiple aliased images that are sampled at different subpixel offsets.
+When someone is holding an object there is a natural and involuntary slight hand movement present.
+Wronski et al.
+shows how this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
 
-To set a wider table, which takes up the whole width of
-the page's live area, use the environment
-\textbf{table*} to enclose the table's contents and
-the table caption, as demonstrated in Table~\ref{tab:typicalCommands} below.  
-As with a single-column table, this wide
-table will ``float" to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table~\ref{tab:typicalCommands} is included in the input file; again, it is
-instructive to compare the placement of the
-table here with the table in the printed dvi
-output of this document.
+Wronski et al.
+analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography.
+They used the rotational measurements from the phone's gyroscope but ignored the effect of translations in the analysis.
+Their analysis showed that hand movement creates uniformly random angular displacements and relatively slow rotation of the device.
 
-\begin{table*}[t]
-\centering
-\caption{Some Typical Commands}
-\label{tab:typicalCommands}
-\begin{tabular}{ccl}
-Command & A Number & Comments \\ \hline
-\texttt{\textbackslash alignauthor}     & 100 & Author alignment\\
-\texttt{\textbackslash numberofauthors} & 200 & Author enumeration\\
-\texttt{\textbackslash table}           & 300 & For tables\\
-\texttt{\textbackslash table*}          & 400 & For wider tables\\ 
-\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
+While hand shake over a long time interval uniformly random, over a short burst it could be more of a straight line.
+Wronski et al. show that this also provides a sufficiently uniform distribution of subpixel samples.
+Using the equidistribution theorem, which states that the sequence $\{a,2a,3a,\dotsc \bmod 1\}$ is uniformly distributed (assuming $a$ is an irrational number), with each pixel as a point sample and a least random scenario (the hand motion is regular and linear), the samples from all frames combined will be approximately uniformly distributed within the subpixel space.
 
-\subsection{Citations}
-\label{sec:citations}
+Wronski et al.
+also tested this concept empirically by measuring subpixel offsets by registration across 20 handheld bursts.
+They found some deviation from a uniform distribution largely caused by pixel locking, which causes a bias towards whole pixel values.
+Overall, the subpixel coverage remained sufficient to be used for super-resolution.
 
-Citations to articles~\cite{Aaronson:2005,Garey:1979,Brun:2008} listed
-in the Bibliography section of your
-article will occur throughout the text of your article.
-You should use BibTeX to automatically produce this bibliography;
-you simply need to insert one of several citation commands with
-a key of the item cited in the proper location in
-the \texttt{.tex} file~\cite{OM:2008}.
-The key is a short reference you invent to uniquely
-identify each work; in this sample document, the key is
-the first author's surname and a
-word from the title.  This identifying key is included
-with each item in the \texttt{.bib} file for your article.
+\subsection{Proposed super-resolution approach}
 
-It is recommended that you precede \texttt{\textbackslash cite} (and 
-\texttt{\textbackslash ref}) commands with a tilde
-character instead of a space, e.g., \texttt{some text\textasciitilde\textbackslash cite}. The tilde gives you a non-breaking space which ensures that your citation won't get
-stranded by itself on the beginning of a line.
+\subsubsection{Kernel reconstruction}
+\label{sec:kernelReconstruction}
 
-The details of the construction of the \texttt{.bib} file
-are beyond the scope of this sample document, but more
-information can be found in the \textit{Author's Guide},
-and exhaustive details in the \textit{\LaTeX\ User's
-Guide}.
+\subsubsection{Motion Robustness}
+\label{sec:robustnessModel}
 
-This article shows only the plainest form
-of the citation command, using \texttt{\textbackslash cite},
-which is all that is needed for our senior seminar.
-You shouldn't use any other forms here.
+\subsection{Results}
 
-\subsection{Theorem-like Constructs}
-\label{sec:theoremLikeConstructs}
+\subsubsection{Limitations}
 
-Other common constructs that may occur in your article are
-the forms for logical constructs like theorems, axioms,
-corollaries and proofs.  There are
-two forms, one produced by the
-command \texttt{\textbackslash newtheorem} and the
-other by the command \texttt{\textbackslash newdef}; perhaps
-the clearest and easiest way to distinguish them is
-to compare the two in the output of this sample document:
+% ----------------------------------------
+\section{Handheld low light photography}
 
-Theorem~\ref{thm:integration} below uses the \textbf{theorem} environment, created by
-the \texttt{\textbackslash newtheorem} command:
+\subsection{Motion Metering}
 
-% You would usually put a \newtheorem command up at the top 
-% of your LaTeX document after the \usepackage commands. It's
-% just here in this example so it's with the text that describes it.
-\newtheorem{theorem}{Theorem}
+\subsection{Motion-adaptive burst merging}
 
-\begin{theorem}
-Let $f$ be continuous on $[a,b]$.  If $G$ is
-an antiderivative for $f$ on $[a,b]$, then
-\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
-\label{thm:integration}
-\end{theorem}
+\subsection{Auto white balance in low-light}
 
-The other uses the \textbf{definition} environment, created
-by the \texttt{\textbackslash newdef} command:
-\newdef{definition}{Definition}
-\begin{definition}
-If $z$ is irrational, then by $e^z$ we mean the
-unique number which has
-logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
-\end{definition}
+\subsection{Tone mapping}
 
-Two lists of constructs that use one of these
-forms is given in the
-\textit{Author's  Guidelines}.
+\subsection{Results}
 
-There is one other similar construct environment, which is
-already set up
-for you; i.e. you must \textit{not} use
-a \texttt{\textbackslash newdef} command to
-create it: the \textbf{proof} environment.  Here
-is a example of its use:
-\begin{proof}
-Suppose on the contrary there exists a real number $L$ such that
-\begin{displaymath}
-\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-\end{displaymath}
-Then
-\begin{align*}
-l &= \lim_{x\rightarrow c} f(x) \\
-  &= \lim_{x\rightarrow c}
-\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ] \\
-  &= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-\frac{f(x)}{g(x)}  \\
-  &= 0\cdot L  \\
-  &= 0,
-\end{align*}
-which contradicts our assumption that $l\neq 0$.
-\end{proof}
+\subsubsection{Limitations}
 
-Complete rules about using these environments and using the
-two different creation commands are in the
-\textit{Author's Guide}; please consult it for more
-detailed instructions.  If you need to use another construct,
-not listed therein, which you want to have the same
-formatting as the Theorem
-or the Definition~\cite{salas:calculus} shown above,
-use the \texttt{\textbackslash newtheorem} or the
-\texttt{\textbackslash newdef} command,
-respectively, to create it.
-
-\subsection*{A {\secit Caveat} for the \TeX\ Expert}
-\label{sec:caveatForExperts}
-
-Because you have just been given permission to
-use the \texttt{\textbackslash newdef} command to create a
-new form, you might think you can
-use \TeX's \texttt{\textbackslash def} to create a
-new command: \textit{Please refrain from doing this!}
-Remember that your \LaTeX\ source code is primarily intended
-to create camera-ready copy, but may be converted
-to other forms -- e.g. HTML. If you inadvertently omit
-some or all of the \texttt{\textbackslash def}s recompilation will
-be, to say the least, problematic.
-
+% ----------------------------------------
 \section{Conclusions}
-\label{sec:conclusions}
-
-This paragraph will end the body of this sample document.
-Remember that you might still have Acknowledgments or
-Appendices; brief samples of these
-follow.  There is still the Bibliography to deal with; and
-we will make a disclaimer about that here: with the exception
-of the reference to the \LaTeX\ book, the citations in
-this paper are to articles which have nothing to
-do with the present subject and are used as
-examples only.
 
@@
 \label{sec:acknowledgments}
-
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.
-
-You want to use the \texttt{\textbackslash section*} version of the \texttt{section}
-command, as an acknowledgments section typically does \emph{not} get
-a number.
-
-It is common (but by no means necessary) for students to thank
-their advisor, and possibly other faculty, friends, and family who provided
-useful feedback on the paper as it was being written.
-
-In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
 

### 2020-09-24 paper.tex
Section Draft update

@@
 
+A bayer filter is a type of color filter array (CFA).
 The majority of digital image sensors in digital cameras and phones use a Bayer filter mosaic pattern to arrange RGB color filters on the sensor.
@@
 
-Background information on demosaicing. Some of this will come from \cite{wiki:Demosaicing}.
+A demosaicing algorithm reconstructs a full color image from the separate red, green, or blue only pixels from an image sensor with a CFA \cite{wiki:Demosaicing}.
+There are many methods for this, the simplest ones interpolate the values for the other two color channels of a given pixel based on nearby pixels from the CFA image of those colors.
 

### 2020-09-24 paper.tex
section draft typo fixes

@@
 
-A bayer filter is a type of color filter array (CFA).
+A Bayer filter is a type of color filter array (CFA).
 The majority of digital image sensors in digital cameras and phones use a Bayer filter mosaic pattern to arrange RGB color filters on the sensor.
@@
 
-Wronski et al.'s approach is a process that starts with the the acquisition of a burst from the continous ring buffer of raw frames in the phone's camera application.
+Wronski et al.'s approach is a process that starts with the the acquisition of a burst from the continuous ring buffer of raw frames in the phone's camera application.
 Next, a single frame is chosen and the rest are aligned to it using a refined version of the algorithm by Hasinoff et al.~\cite{Hasinoff2016}.

### 2020-10-09 paper.tex
4 page draft

@@
 \usepackage{url}
+
+\usepackage{subcaption}
+
+\usepackage{graphicx}
+
+\graphicspath{ {./figures/} }
 
@@
 \begin{abstract}
-(This is currently just a topic statement more than an abstract)
 
-My topic is recent advances in the computational photography used in smartphones.
-Specifically looking at the use of multi-frame algorithms to increase image resolution and quality.
-I'm mainly looking at a couple of papers by researchers at Google on algorithms used in the Super-Res Zoom and Night Sight features of the Google Pixel series of smartphones.
+Smartphone cameras present many challenges, most of which come from the need for them to be physically small.
+Their small size puts a fundamental limit on their ability to resolve detail and especially affects the ability to zoom and low-light photography.
+This paper presents two approaches to improve smartphone photography through software techniques.
+The first is handheld super-resolution which uses natural hand movement to improve the resolution of burst images.
+The second approach is a system which improves low light photography in smartphones.
+
 \end{abstract}
@@
 
-Smartphone cameras present many challenges, most of which come from the need for them to be physically small.
-Their small size puts a fundamental limit on their ability to resolve detail and especially affects low-light photography.
-This paper will look at two approaches to improve smartphone photography through software techniques.
+Smartphone cameras use very small sensors with fixed apertures.
+This means their ability to gather light is significantly reduced compared to larger dedicated cameras.
+Most modern smartphones use a burst system to capture multiple images when the user presses the capture button and merges them to improve image quality.
+
+In this paper, I first introduce some background information concerning various photography and signal processing terms.
+I then introduce and discuss two approaches to improving smartphone photography that augment the existing burst pipeline.
+The first uses hand movement to increase spatial resolution (Section~\ref{sec:handheldSuperRes}).
+The second approach is a a system that uses a number of techniques to improve the low-light abilities of smartphone cameras (Section~\ref{sec:handheldLowLight}.
 
@@
 \section{Background}
+\label{sec:background}
+
+In this section I explain several key concepts that are important for the work in Sections \ref{sec:handheldSuperRes} and \ref{sec:handheldLowLight}.
+This includes Burst photography (Section~\ref{sec:background:burstPhotography}), Bayer filters (Section~\ref{sec:background:bayerFilter}), demosaicing (Section~\ref{sec:background:demosaicing}), aliasing (Section~\ref{sec:background:aliasing}), super-resolution (Section~\ref{sec:background:superResolution}), and kernels (Section~\ref{sec:background:kernels}).
 
 \subsection{Burst photography}
+\label{sec:background:burstPhotography}
 
@@
 Generally, burst processing involves taking a series of raw exposures and merging them together to form the final image.
-Most smartphones operate in a \emph{Zero-Shutter Lag} mode by default.
-In this mode raw frames are continuously captured to a ring buffer while the camera app is open.
-When the user presses the shutter button, a number of the most recent frames are sent to the camera processing pipeline.
+Most smartphones operate in a \emph{zero-shutter lag} mode by default.
+In this mode raw frames (the full unprocessed sensor output) are continuously captured to a temporary memory while the camera app is open.
+When the user presses the shutter button, several of the most recent frames are sent to the camera processing pipeline.
 
@@
 \cite{Hasinoff2016} which used bursts of constant low-exposure frames to increase dynamic range and signal-to-noise ratio.
-% TODO: more on Hasinoff
-
-\newpage
+\todo[inline]{more on Hasinoff}
 
 \subsection{Bayer filter}
+\label{sec:background:bayerFilter}
 
@@
 The majority of digital image sensors in digital cameras and phones use a Bayer filter mosaic pattern to arrange RGB color filters on the sensor.
-The pattern consists of 50\% green, 25\% red, and 25\% blue pixels \cite{wiki:BayerFilter}.
+The pattern consists of 50\% green, 25\% red, and 25\% blue pixels.
 This ratio emulates the color sensitivity of the human eye.
-Due to this pattern on the sensor, the raw output of a digital camera also has each pixel filtered to only red, green, or blue and a demosaicing algorithm must be used to interpolate the other values for each pixel \cite{wiki:BayerFilter}.
+Due to this pattern on the sensor, the raw output of a digital camera also has each pixel filtered to only red, green, or blue and a demosaicing algorithm must be used to interpolate the other values for each pixel.
+\cite{wiki:BayerFilter}
 
@@
 \centering
-\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
+%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
+\includegraphics[width=3in]{Bayer_pattern_on_sensor}
 \caption{A Bayer pattern on a sensor in isometric perspective \cite{wiki:BayerFilter}.}
@@
 \subsection{Demosaicing}
+\label{sec:background:demosaicing}
 
@@
 The demosaicing process may introduce various artifacts in the final image. These typically include false color artifacts like zippering and Moiré patterns \cite{Wronski2019}.
+\todo[inline]{I might need to explain zippering and Moiré patterns}
 
 \subsection{Aliasing}
+\label{sec:background:aliasing}
 
-Background information on aliasing, specifically in regard to signal processing. Some of this will come from \cite{wiki:Aliasing}.
+\todo[inline]{Background information on aliasing, specifically in regard to signal processing. Some of this will come from \cite{wiki:Aliasing}.}
 
 \subsection{Super-resolution}
+\label{sec:background:superResolution}
+
+Generally, super-resolution is a type of technique which increases the resolution of an image.
+While there are techniques that work on a single image, we will focus on using multiple frames for super-resolution.
+The main requirements of multi-frame super-resolution have been identified as accurate sub-pixel registration, the existence of aliasing, and a good signal-to-noise ratio \cite{Wronski2019}.
+Sub-pixel registration refers to the ability to align frames with sub-pixel accuracy, that is
+
+%requirements
+%- subpixels
+%- aliasing
+
+\todo[inline]{subpixel offsets}
+
+\subsection{Kernels}
+\label{sec:background:kernels}
 
@@
 \section{Handheld super-resolution}
+\label{sec:handheldSuperRes}
+
+\todo[inline]{actually mention how this is used to allow better zooming}
 
@@
 % TODO: Size correctly for width
-\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
+%\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
+\includegraphics[width=7in]{Wronski2019-figure-2}
 \caption{An overview of the approach used by Wronski et al. \cite{Wronski2019}}
@@
 % TODO: elaborate on the alignment algorithm
-Each frame's local contributions are estimated through kernel regression (Section~\ref{sec:kernelReconstruction}) and accumulated across an whole burst for each of the three color planes.
+Each frame's local contributions are estimated through kernel regression (Section~\ref{sec:kernelReconstruction}) and accumulated across a whole burst for each of the three color planes.
 Next, in parallel, the kernel shapes are adjusted based on estimated local gradients and the sample contributions are adjusted weighted based on a robustness model (Section~\ref{sec:robustnessModel}).
@@
 \subsection{Hand movement based super-resolution}
+
+\todo[inline]{use tree figure from Wronski either here or in background about super res/subpixels}
 
@@
 
-Wronski et al.
-analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography.
-They used the rotational measurements from the phone's gyroscope but ignored the effect of translations in the analysis.
+Wronski et al. analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography.
+%They used the rotational measurements from the phone's gyroscope but ignored the effect of translations in the analysis.
 Their analysis showed that hand movement creates uniformly random angular displacements and relatively slow rotation of the device.
 
-While hand shake over a long time interval uniformly random, over a short burst it could be more of a straight line.
+While hand shake over a long time is interval uniformly random, over a short burst it could be more of a straight line.
 Wronski et al. show that this also provides a sufficiently uniform distribution of subpixel samples.
@@
 
+Super-resolution involves recreating a high resolution signal from many lower resolution samples.
+Due to the random distribution of pixel shifts that result from hand movement, kernel regression was deemed a good technique for this reconstruction.
+Wronski et al.
+used anisotropic Gaussian Radial Basis Function (RBF) kernels (Section~\ref{sec:kernelReconstruction}) and a statistical robustness model (Section~\ref{sec:robustnessModel}) to combine burst frames into a higher resolution image.
+
 \subsubsection{Kernel reconstruction}
@@
 
+The core idea of Wronski et al's algorithm is considering the pixels of multiple raw frames as randomly offset, aliased, and noisy measurements of three original continuous signals, one for each color channel.
+The algorithm creates the final output image pixel-by-pixel.
+For each output pixel, it evaluates the local contributions to each of the three color channels from different input frames since each raw image pixel is specific to a single color channel.
+
+The process for each color channel can be described as:
+\begin{equation}
+C(x, y)=\frac{\sum_{n} \sum_{i} c_{n, i} \cdot w_{n, i} \cdot \hat{R}_{n}}{\sum_{n} \sum_{i} w_{n, i} \cdot \hat{R}_{n}}
+\end{equation}
+where $(x,y)$ is the coordinates of the pixel, $\sum_{n}$ is the sum of values of that pixel over all contributing frames, $\sum_{i}$ is the sum over samples (pixels) in a local neighborhood ($3\times 3$), $c_{n, i}$ is the value of the pixel at frame $n$ and sample $i$, $w_{n, i}$ is the local sample weight, and $\hat{R}_{n}$ is the local motion robustness score (Section~\ref{sec:robustnessModel}), it is equal to 1 for the base frame since that's what the other frames are aligned against.
+\cite{Wronski2019}
+\todo[inline]{don't know if i actually need this equation to explain the idea}
+
+The local pixel weights are calculated using a 2D unnormalized anisotropic Gaussian radial basis function kernel:
+\begin{equation}
+w_{n, i}=\exp \left(-\frac{1}{2} d_{i}^{T} \Omega^{-1} d_{i}\right)
+\end{equation}
+Where $\Omega$ is the kernel covariance matrix and $d_{i}$ is the offset vector of sample i to the output pixel ($d_{i}=\left[x_{i}-x_{0}, y_{i}-y_{0}\right]^{T}$).
+\todo[inline]{needs more on this function, what things are and why}
+
+The kernel covariance matrix is computed by analyzing the local gradient structure tensor of every frame. This is done on half-resolution images to improve performance and resistance to noise.
+\todo[inline]{how much detail do I need on this, the original paper spends half a page on the specifics}
+
 \subsubsection{Motion Robustness}
 \label{sec:robustnessModel}
+
+It is difficult to reliably align the sequence of images in a burst and even assuming a perfect alignment, changes in the scene and occlusion would still result in some areas of the scene being poorly represented in many frames of the burst.
+This needs to be taken into account to prevent severe artifacting.
+To combine frames robustly, a confidence level is assigned to the local neighborhood of each pixel, the map of these confidences is called a \emph{robustness map}.
+% try to reword below
+A robustness value of one corresponds to fully merged areas while a value of zero means rejected areas.
+
+The robustness values are determined by a statistical robustness model which must distinguish between aliasing, as it is needed for super-resolution, and frame misalignment, which is detrimental to super-resolution.
+Wronski et al.
+observed that ``areas prone to aliasing have large spatial variance even within a single frame" \cite{Wronski2019}.
+
+\todo[inline]{I don't know how much detail to go into the motion robustness algorithm}
 
@@
 \section{Handheld low light photography}
+\label{sec:handheldLowLight}
+
+Liba et al.
+\cite{Liba2019} introduce a system for capturing photos in very low light that produces improved color and less noise.
+The system uses ``motion metering" (Section~\ref{sec:motionMetering}) to enable motion aware burst merging (Section~\ref{sec:motionBurstMerging}).
+They also use a learning-based auto white balance algorithm (Section~\ref{sec:autoWhiteBal}) and tone mapping (Section~\ref{sec:toneMapping}) to produce better colors in low light.
+
+% The second primary source is ``Handheld Mobile Photography in Very Low Light" \cite{Liba2019}.
+This paper describes a new photography system (or pipeline) that improves the ability to capture photos in low light.
+This is done mainly through improvements in motion metering, motion-aware burst merging, and low-light optimized auto white balance.
+This system also builds on work done in \cite{Hasinoff2016}.
 
 \subsection{Motion Metering}
+\label{sec:motionMetering}
+
+When capturing a burst of images on a smartphone, the exposure time and gain (ISO) needs to be selected for each frame.
+Liba et al.
+use the same strategy as in \cite{Hasinoff2016} and capture all frames in the burst with the same exposure time and ISO.
+For effective low-light photography these settings need to be automatically selected within a set of constraints.
+
+The ``motion metering" described by Liba et al.
+selects the exposure time and gain by predicting future motion in the scene and motion of the camera itself.
+This is used to select slower exposures for scenes with no motion and shorter exposures for those with motion.
+
+\subsubsection{Motion prediction}
+
+\subsubsection{Stability detection}
+
+More experienced photographers often brace their device against a surface or put it on a tripod in low-light situations.
+If this can be detected, longer exposures can be used without risk of motion blur.
+Liba et al.'s motion metering system captures up to 333 ms exposures when handheld but when stabilized it captures up to 1 second exposures.
+To detect a stabilized device, measurements from the device's gyroscope are averaged over a short time, 1.466 s, before the shutter button was pressed while masking out the last 0.4 s of the measurements to prevent vibrations from the user pressing the shutter button from affecting the result.
+
+\subsubsection{Exposure time selection}
+
+\begin{figure}
+\centering
+\includegraphics[width=3in]{figures/liba2019-figure-7.pdf}
+\caption{ \cite{Liba2019}}
+\label{fig:BayerPattern}
+
+% Cameras typically encode the exposure time vs. gain tradeoff with
+% a static “exposure schedule”, depicted by the thicker blue line. If the ex-
+% posure schedule curve is higher, the camera favors higher SNR and more
+% motion blur, and if the curve is lower, the camera favors lower SNR and
+% less motion blur. Instead of a static schedule, we use a dynamic schedule
+% based on the measured scene motion. This allows the schedule to provide a
+% better tradeoff for the scene’s conditions. Zero motion (orange line) leads
+% to the slowest possible schedule, and as motion increases, the schedule
+% favors proportionally faster exposure times. The flat regions of the dynamic
+% exposure schedules denote the “blur-limiting exposure time" in Equation 6
+
+\end{figure}
 
 \subsection{Motion-adaptive burst merging}
+\label{sec:motionBurstMerging}
+
+For burst merging, Liba et al.
+built upon the Fourier domain temporal merging from Hasinoff et al.
+\cite{Hasinoff2016}.
+This is used on some devices while those with a fast enough system-on-chip use the merging method from Wronski et al.
+(Section~\ref{sec:handheldSuperRes}).
+
+\todo[inline]{don't know how much I need to talk about this if at all considering on modern phones the Wronski method is used anyway}
 
 \subsection{Auto white balance in low-light}
+\label{sec:autoWhiteBal}
+
+The Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image.
+The goal of AWB is to determine the color of the majority of the illumination in the scene and correct the colors in the image such that they appear to be lit by a neutral illumination.
+%This is closely related to color constancy
+
+\todo[inline]{more}
 
 \subsection{Tone mapping}
+\label{sec:toneMapping}
+
+Tone mapping is the process of mapping colors from a high-dynamic-range image to a medium with a more limited dynamic range \cite{wiki:ToneMapping}.
+This is usually done by applying what are called tone mapping operators (TMOs).
+Some TMOs attempt to create results close to human vision, while others produce a more artistic rendition.
+Human vision loses color sensitivity as light levels are reduced.
+Liba et al.
+developed a TMO that maintains vibrant color in dark scenes without looking artificial with inspiration from Artistic techniques to maintain a nighttime aesthetic.
+Their TMO uses various heuristics on top of the tone mapping of Hasinoff et al.
+\cite{Hasinoff2016}.
+These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.
+\cite{Liba2019}
+
+\todo[inline]{explain these heuristics in more detail}
+
+\begin{figure}
+\centering
+%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
+\begin{subfigure}{1in}
+\centering
+\includegraphics[width=1in]{figures/liba2019-figure-14a-95quality.jpg}
+\caption{Baseline}
+\label{fig:toneMapping:baseline}
+\end{subfigure}
+\begin{subfigure}{1in}
+\centering
+\includegraphics[width=1in]{figures/liba2019-figure-14b-95quality.jpg}
+\caption{CLAHE}
+\label{fig:toneMapping:clahe}
+\end{subfigure}
+\begin{subfigure}{1in}
+\centering
+\includegraphics[width=1in]{figures/liba2019-figure-14c-95quality.jpg}
+\caption{Liba et al.}
+\label{fig:toneMapping:liba}
+\end{subfigure}
+
+\caption{An example of tone mapping \cite{Liba2019}}
+\todo[inline]{needs better caption}
+% A nighttime scene (about 0.4 lux) that demonstrates the difficulty
+% in achieving balanced tonal quality. (a) The tone mapping of [Hasinoff et al .
+% 2016] results in most of the image being too dark and is not a pleasing
+% photograph. (b) Applying a tone mapping technique that brightens the
+% image using histogram equalization (CLAHE [Zuiderveld 1994]) reveals
+% more detail, but the photo lacks global contrast. (c) The photo with our tone
+% mapping retains the details of (b), but contains more global contrast and
+% enough dark areas to show that the captured scene is dark.
+
+\label{fig:toneMapping}
+\end{figure}
 

### 2020-10-26 paper.tex
Complete First Draft

@@
 
-\title{Recent advances in smartphone computational photography}
+\title{Recent Advances in Smartphone Computational Photography}
 
@@
 The first uses hand movement to increase spatial resolution (Section~\ref{sec:handheldSuperRes}).
-The second approach is a a system that uses a number of techniques to improve the low-light abilities of smartphone cameras (Section~\ref{sec:handheldLowLight}.
+The second approach is a a system that uses a number of techniques to improve the low-light abilities of smartphone cameras (Section~\ref{sec:handheldLowLight}).
 
@@
 
+An image processing pipeline refers to the process the raw data from a camera sensor goes through to be turned into the final image file that can be displayed and shared.
+
 Most smartphones use a burst processing pipeline for their cameras.
@@
 \cite{Hasinoff2016} which used bursts of constant low-exposure frames to increase dynamic range and signal-to-noise ratio.
-\todo[inline]{more on Hasinoff}
+
+%\todo[inline]{more on Hasinoff}
 
@@
 The pattern consists of 50\% green, 25\% red, and 25\% blue pixels.
-This ratio emulates the color sensitivity of the human eye.
-Due to this pattern on the sensor, the raw output of a digital camera also has each pixel filtered to only red, green, or blue and a demosaicing algorithm must be used to interpolate the other values for each pixel.
-\cite{wiki:BayerFilter}
+This ratio emulates the color sensitivity of the human eye, which is more sensitive to green than it is to blue and red.
+Due to this pattern on the sensor, the raw output of a digital camera also has each pixel filtered to only red, green, or blue and a demosaicing algorithm must be used to interpolate the other values for each pixel.~\cite{wiki:BayerFilter}
 
@@
 \includegraphics[width=3in]{Bayer_pattern_on_sensor}
-\caption{A Bayer pattern on a sensor in isometric perspective \cite{wiki:BayerFilter}.}
+\caption{A Bayer pattern on a sensor in isometric perspective.~\cite{wiki:BayerFilter}}
 \label{fig:BayerPattern}
 \end{figure}
+
+\begin{figure*}[t]
+\centering
+% TODO: Size correctly for width
+%\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
+\includegraphics[width=7in]{Wronski2019-figure-2}
+\caption{An overview of the approach used by Wronski et al.~\cite{Wronski2019}.
+The initial burst of input frames (a) are aligned (d) to a base frame.
+The local features in the frames (b) are used to create kernels (c) (Section~\ref{sec:kernelReconstruction}) which are used along with the motion robustness model (f) (Section~\ref{sec:robustnessModel}) to combine (g) the frames separately for each color channel.
+The final image (h) is produced by normalizing the results of each channel.}
+\label{fig:Wronski2019Fig2}
+\end{figure*}
 
@@
 
-A demosaicing algorithm reconstructs a full color image from the separate red, green, or blue only pixels from an image sensor with a CFA \cite{wiki:Demosaicing}.
-There are many methods for this, the simplest ones interpolate the values for the other two color channels of a given pixel based on nearby pixels from the CFA image of those colors.
+A demosaicing algorithm reconstructs a full color image from the separate pixels that have been filtered with a CFA to just one channel: red, green, or blue \cite{wiki:Demosaicing}.
+There are many methods for this; the simplest ones interpolate the values for the other two color channels of a given pixel based on nearby pixels from the CFA image of those colors.
 
-The demosaicing process may introduce various artifacts in the final image. These typically include false color artifacts like zippering and Moiré patterns \cite{Wronski2019}.
-\todo[inline]{I might need to explain zippering and Moiré patterns}
+This process means two-thirds of the final image is reconstructed from the available data.
+The demosaicing process may introduce various artifacts in the final image due to aliasing.~\cite{blog:Wronski2018}
+%These typically include false color artifacts like zippering and Moiré patterns~\cite{Wronski2019}.
+%\todo[inline]{I might need to explain zippering and Moiré patterns}
 
@@
 
-\todo[inline]{Background information on aliasing, specifically in regard to signal processing. Some of this will come from \cite{wiki:Aliasing}.}
+Aliasing is an effect that happens when the camera sensor is unable to correctly represent the patterns and details present in a scene due to its resolution.
+This often results in artifacts such as Moiré patterns.
+\cite{blog:Wronski2018}.
 
@@
 While there are techniques that work on a single image, we will focus on using multiple frames for super-resolution.
-The main requirements of multi-frame super-resolution have been identified as accurate sub-pixel registration, the existence of aliasing, and a good signal-to-noise ratio \cite{Wronski2019}.
-Sub-pixel registration refers to the ability to align frames with sub-pixel accuracy, that is
-
-%requirements
-%- subpixels
-%- aliasing
-
-\todo[inline]{subpixel offsets}
+The main requirements of multi-frame super-resolution are aliased input frames that are sampled at different subpixel offsets (the individual pixel points in each frame captures a different sample of the area it represents)~\cite{Wronski2019}.
 
@@
 \label{sec:background:kernels}
+
+\todo[inline]{TODO: Explain kernels either here or in later section where it comes up}
 
@@
 
-\todo[inline]{actually mention how this is used to allow better zooming}
-
-\begin{figure*}[t!]
-\centering
-% TODO: Size correctly for width
-%\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
-\includegraphics[width=7in]{Wronski2019-figure-2}
-\caption{An overview of the approach used by Wronski et al. \cite{Wronski2019}}
-\label{fig:Wronski2019Fig2}
-\end{figure*}
+%\todo[inline]{actually mention how this is used to allow better zooming}
 
@@
 It uses natural hand motion and is efficient enough to work in the background on smartphones.
+This algorithm is used as the merging algorithm in the camera pipeline of the Google Pixel 3 and newer and is what allows for the Pixel's ``super-res zoom" feature.
 
@@
 
-Wronski et al.'s approach is a process that starts with the the acquisition of a burst from the continuous ring buffer of raw frames in the phone's camera application.
+% The algorithm tackles the tasks of demosaicing and super-
+% resolution jointly and formulates the problem as the reconstruction
+% and interpolation of a continuous signal from a set of sparse samples.
+
+The approach by Wronski et al.~\cite{Wronski2019}, as shown in Figure~\ref{fig:Wronski2019Fig2}, is a process that starts with the the acquisition of a burst from the continuous ring buffer of raw frames in the phone's camera application.
 Next, a single frame is chosen and the rest are aligned to it using a refined version of the algorithm by Hasinoff et al.~\cite{Hasinoff2016}.
@@
 Each frame's local contributions are estimated through kernel regression (Section~\ref{sec:kernelReconstruction}) and accumulated across a whole burst for each of the three color planes.
-Next, in parallel, the kernel shapes are adjusted based on estimated local gradients and the sample contributions are adjusted weighted based on a robustness model (Section~\ref{sec:robustnessModel}).
-The final RGB image is obtained by normalizing the accumulated contributions for each of the three color planes and merging them together.
+This involves the kernel shapes being adjusted based on estimated local gradients and, at the same time, the sample contributions are adjusted weighted based on a statistical robustness model (Section~\ref{sec:robustnessModel}).
+
+The final RGB image is obtained by normalizing the accumulated contributions for each of the three color planes and merging them together. This can then be sent to the rest of the imaging pipeline.
 
@@
 
-\todo[inline]{use tree figure from Wronski either here or in background about super res/subpixels}
+\begin{figure*}
+\centering
+% TODO: Size correctly for width
+\includegraphics[width=7in]{Wronski2019-figure-4}
+\caption{An illustration of subpixel displacements from a burst of four frames with linear hand motion.
+Each frame is offset by half a pixel on the x-axis and a quarter pixel on the y-axis from the last frame.
+After alignment, the pixel centers (black dots) uniformly cover the image with greater density than a single frame.~\cite{Wronski2019}}
+\label{fig:Wronski2019Fig4}
 
-One of the important conditions for multi-frame super-resolution is that the input contains multiple aliased images that are sampled at different subpixel offsets.
+% Illustration of a burst of four frames with linear hand motion. Each frame is offset from the
+% previous frame by half a pixel along the x-axis and a quarter pixel along the y-axis due to the hand motion. Afuer alignment to the base frame, the pixel
+% centers (black dots) uniformly cover the resampling grid (grey lines) at an increased density. In practice, the distribution is more random than in this simplified
+% example.
+
+\end{figure*}
+
+One of the important conditions for multi-frame super-resolution is that the input contains multiple images that are sampled at different subpixel offsets.
 When someone is holding an object there is a natural and involuntary slight hand movement present.
 Wronski et al.
-shows how this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
+show how this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
 
@@
 
-While hand shake over a long time is interval uniformly random, over a short burst it could be more of a straight line.
+While hand shake over a long time is uniformly random, over a short burst it could be more of a straight line.
 Wronski et al. show that this also provides a sufficiently uniform distribution of subpixel samples.
-Using the equidistribution theorem, which states that the sequence $\{a,2a,3a,\dotsc \bmod 1\}$ is uniformly distributed (assuming $a$ is an irrational number), with each pixel as a point sample and a least random scenario (the hand motion is regular and linear), the samples from all frames combined will be approximately uniformly distributed within the subpixel space.
+Using the equidistribution theorem, which states that the sequence $\{a,2a,3a,\dotsc \bmod 1\}$ is uniformly distributed (assuming $a$ is an irrational number), with each pixel as a point sample and a least random scenario (the hand motion is regular and linear), the samples from all frames combined will be approximately uniformly distributed within the subpixel space (Figure~\ref{fig:Wronski2019Fig4}).
 
@@
 
-Super-resolution involves recreating a high resolution signal from many lower resolution samples.
+Super-resolution involves recreating a higher resolution signal from many lower resolution samples.
 Due to the random distribution of pixel shifts that result from hand movement, kernel regression was deemed a good technique for this reconstruction.
@@
 
-The core idea of Wronski et al's algorithm is considering the pixels of multiple raw frames as randomly offset, aliased, and noisy measurements of three original continuous signals, one for each color channel.
+The core idea of the algorithm is considering the pixels of multiple raw frames as randomly offset, aliased, and noisy measurements of three original continuous signals, one for each color channel.
 The algorithm creates the final output image pixel-by-pixel.
@@
 
-The process for each color channel can be described as:
+The process for each pixel in each color channel can be described as:
 \begin{equation}
@@
 \end{equation}
-where $(x,y)$ is the coordinates of the pixel, $\sum_{n}$ is the sum of values of that pixel over all contributing frames, $\sum_{i}$ is the sum over samples (pixels) in a local neighborhood ($3\times 3$), $c_{n, i}$ is the value of the pixel at frame $n$ and sample $i$, $w_{n, i}$ is the local sample weight, and $\hat{R}_{n}$ is the local motion robustness score (Section~\ref{sec:robustnessModel}), it is equal to 1 for the base frame since that's what the other frames are aligned against.
-\cite{Wronski2019}
-\todo[inline]{don't know if i actually need this equation to explain the idea}
 
-The local pixel weights are calculated using a 2D unnormalized anisotropic Gaussian radial basis function kernel:
-\begin{equation}
-w_{n, i}=\exp \left(-\frac{1}{2} d_{i}^{T} \Omega^{-1} d_{i}\right)
-\end{equation}
-Where $\Omega$ is the kernel covariance matrix and $d_{i}$ is the offset vector of sample i to the output pixel ($d_{i}=\left[x_{i}-x_{0}, y_{i}-y_{0}\right]^{T}$).
-\todo[inline]{needs more on this function, what things are and why}
+\begin{description}
+  \item[$(x,y)$] refers to the coordinates of the pixel. 
+  \item[$\sum_{n}$] is a sum over all contributing frames.
+  \item[$\sum_{i}$] is the sum over samples (pixels) in a local $3\times 3$ neighborhood centered on the target pixel.
+  \item[$c_{n, i}$] is the value of the individual pixel at frame $n$ and sample $i$.
+  \item[$w_{n, i}$] is the local sample weight for the pixel at frame $n$ and sample $i$, which is described below. 
+  \item[$\hat{R}_{n}$] is the local motion robustness score at $(x,y)$ described in Section~\ref{sec:robustnessModel}.
+\end{description}
 
-The kernel covariance matrix is computed by analyzing the local gradient structure tensor of every frame. This is done on half-resolution images to improve performance and resistance to noise.
-\todo[inline]{how much detail do I need on this, the original paper spends half a page on the specifics}
+%$(x,y)$ refers to the coordinates of the pixel. 
+%$\sum_{n}$ is a sum over all contributing frames. 
+%\sum_{i}$ is the sum over samples (pixels) in a local $3\times 3$ neighborhood.
+
+%$c_{n, i}$ is the value of the pixel at frame $n$ and sample $i$. 
+%$w_{n, i}$ is the local sample weight, which is described below. 
+%$\hat{R}_{n}$ is the local motion robustness score  described in Section~\ref{sec:robustnessModel}. \cite{Wronski2019}
+
+\begin{figure}
+\centering
+\includegraphics[width=3in]{figures/Wronski2019-figure-8.pdf}
+\caption{Plots of relative weights as a function of local features~\cite{Wronski2019}}
+\label{fig:Wronski2019Fig8}
+% Merge kernels: Plots of relative weights in difgerent 3 × 3 sampling
+% kernels as a function of local tensor features
+\end{figure}
+
+The local pixel weights ($w_{n, i}$) come from a 2D unnormalized anisotropic Gaussian radial basis function kernel which is calculated for the $3\times 3$ neighborhood around the pixel at $(x,y)$.
+It produces a $3\times 3$ matrix of values between 0 and 1 taking into account edges and sharp features in the local area.
+These weights determine how much each pixel in the $3\times 3$ area will contribute to the target pixel for that frame.
+An example of the matrices that result from the kernel function can be seen in Figure~\ref{fig:Wronski2019Fig8}.
+
+%\begin{equation}
+%w_{n, i}=\exp \left(-\frac{1}{2} d_{i}^{T} \Omega^{-1} d_{i}\right)
+%\end{equation}
+%Where $\Omega$ is the kernel covariance matrix and $d_{i}$ is the offset vector of sample i to the output pixel ($d_{i}=\left[x_{i}-x_{0}, y_{i}-y_{0}\right]^{T}$).
+%\todo[inline]{needs more on this function, what things are and why}
+
+%The kernel covariance matrix is computed by analyzing the local gradient structure tensor of every frame. This is done on half-resolution images to improve performance and resistance to noise.
+%\todo[inline]{how much detail do I need on this, the original paper spends half a page on the specifics}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=3in]{figures/wronski2019-figure-9-95quality.jpg}
+\caption{A photograph of a moving bus: \textbf{Left}: Without a motion robustness model there are alignment errors and occlusions that result in tiling and ghosting artifacts.
+\textbf{Middle}: The robustness mask produced by the robustness model.
+White regions are those with all frames contributing to the final merged image and darker regions are those with a lower number of contributing frames.
+\textbf{Right}: The result of using the robustness model when merging frames.
+\cite{Wronski2019}}
+\label{fig:Wronski2019Fig9}
+\end{figure}
+
 It is difficult to reliably align the sequence of images in a burst and even assuming a perfect alignment, changes in the scene and occlusion would still result in some areas of the scene being poorly represented in many frames of the burst.
 This needs to be taken into account to prevent severe artifacting.
-To combine frames robustly, a confidence level is assigned to the local neighborhood of each pixel, the map of these confidences is called a \emph{robustness map}.
+To combine frames robustly, a confidence level is assigned to the local neighborhood of each pixel; the map of these confidences is called a \emph{robustness map}.
 % try to reword below
-A robustness value of one corresponds to fully merged areas while a value of zero means rejected areas.
+A robustness value of one corresponds to fully merged areas while a value of zero means rejected areas. The base frame is assigned a value of 1 because it is what the other frames are aligned to.
 
@@
 Wronski et al.
-observed that ``areas prone to aliasing have large spatial variance even within a single frame" \cite{Wronski2019}.
+observed that ``areas prone to aliasing have large spatial variance even within a single frame"~\cite{Wronski2019}.
 
-\todo[inline]{I don't know how much detail to go into the motion robustness algorithm}
+\todo[inline]{TODO: I need to explain specifics of how the motion robustness model works}
 
 \subsection{Results}
+
+\todo[inline]{TODO: Results and limitations}
 
@@
 \cite{Liba2019} introduce a system for capturing photos in very low light that produces improved color and less noise.
-The system uses ``motion metering" (Section~\ref{sec:motionMetering}) to enable motion aware burst merging (Section~\ref{sec:motionBurstMerging}).
+This system is used for the ``Night Sight" feature on Google Pixel phones.
+The system builds on the existing burst pipeline (from Hasinoff et al.~\cite{Hasinoff2016}) and uses the burst merging algorithm from Wronski et al.~\cite{Wronski2019} on Pixel 3 and newer with modifications to improve low-light photography.
+It uses a \emph{positive-shutter-lag} mode rather than the \emph{zero-shutter-lag} mode used for normal photography to allow for longer exposures.
+
+The main improvements to the camera pipeline are the use of ``motion metering (Section~\ref{sec:motionMetering}) to calculate the settings for each frame based on predicted motion in the scene and camera, a learning-based low-light optimized auto white balance algorithm (Section~\ref{sec:autoWhiteBal}), and tone mapping (Section~\ref{sec:toneMapping}) to produce better colors in low light.
+
+%The system uses ``motion metering" (Section~\ref{sec:motionMetering}) to enable motion aware burst merging (Section~\ref{sec:motionBurstMerging}).
 They also use a learning-based auto white balance algorithm (Section~\ref{sec:autoWhiteBal}) and tone mapping (Section~\ref{sec:toneMapping}) to produce better colors in low light.
@@
 
-When capturing a burst of images on a smartphone, the exposure time and gain (ISO) needs to be selected for each frame.
+When capturing a burst of images on a smartphone, the exposure time and sensor gain (ISO) needs to be selected for each frame.
 Liba et al.
-use the same strategy as in \cite{Hasinoff2016} and capture all frames in the burst with the same exposure time and ISO.
-For effective low-light photography these settings need to be automatically selected within a set of constraints.
+use the same strategy as in~\cite{Hasinoff2016} and capture all frames in the burst with the same exposure time and ISO.
+For effective low-light photography these settings need to be automatically selected within the constraints of keeping the total capture time low ($\leq$ 6 seconds) and the total number of frames within the device's memory limits.
+
+This process consists of splitting the target sensitivity, which is based on the brightness of the scene, into exposure time and gain and calculating the number of frames with respect to the time and memory limits.
+The traditional method used in Hasinoff et al.~\cite{Hasinoff2016} uses a fixed ``exposure schedule" that simply keeps the exposure time low to limit motion blur.
+This method usually works well but can be improved for low-light photography.
 
@@
 selects the exposure time and gain by predicting future motion in the scene and motion of the camera itself.
-This is used to select slower exposures for scenes with no motion and shorter exposures for those with motion.
+This is used to select slower exposures for scenes with no motion and shorter exposures for those with motion to reduce motion blur when needed and increase signal-to-noise ration when possible.
 
-\subsubsection{Motion prediction}
+%\subsubsection{Motion prediction}
 
@@
 
-\subsubsection{Exposure time selection}
+%\subsubsection{Exposure time selection}
 
@@
 \includegraphics[width=3in]{figures/liba2019-figure-7.pdf}
-\caption{ \cite{Liba2019}}
-\label{fig:BayerPattern}
+\caption{A traditional static exposure schedule compared to the dynamic exposure scheudle of Liba et al. at various levels of motion.~\cite{Liba2019}}
+\label{fig:exposure}
 
@@
 
-\subsection{Motion-adaptive burst merging}
-\label{sec:motionBurstMerging}
+%\subsection{Motion-adaptive burst merging}
+%\label{sec:motionBurstMerging}
 
-For burst merging, Liba et al.
+%For burst merging, Liba et al.
 built upon the Fourier domain temporal merging from Hasinoff et al.
@@
 
-\todo[inline]{don't know how much I need to talk about this if at all considering on modern phones the Wronski method is used anyway}
+%\todo[inline]{don't know how much I need to talk about this if at all considering on modern phones the Wronski method is used anyway}
 
@@
 
-The Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image.
-The goal of AWB is to determine the color of the majority of the illumination in the scene and correct the colors in the image such that they appear to be lit by a neutral illumination.
-%This is closely related to color constancy
+Humans perceive color correctly of things even when lit with colored light, a phenomenon called color constancy.
+This perception can break down when a photograph is taken under one type of light and viewed under different light.
+It will look tinted (Figure~\ref{fig:whitebalance:default}).
+Cameras correct for this by determining the color of the majority of the illumination in the scene and correcting the colors in the image such that they appear to be lit by a neutral (white) illumination.
+This Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image. \cite{Liba2019, blog:Levoy2018}
 
-\todo[inline]{more}
+The current best-performing color constancy algorithm is ``Fast Fourier Color Constancy" (FFCC) \cite{Barron2017}.
+It is a machine learning based algorithm that ``works by learning and then applying a filter to a histogram of log-chrominance pixel and edge intensities"~\cite{Liba2019}.
+Liba et al.
+trained the FFCC algorithm with a new dataset and error metric to better handle challenging low-light scenes (Figure~\ref{fig:whitebalance}).
+
+Liba et al.
+collected 5000 images using mobile devices of various scenes that demonstrate a range of light levels to be able to train their implementation of FFCC.
+To obtain better ground truth coloring for the training data they had real professional photographers manually white balance the images with the most ``aesthetically preferable" white balance for the scene rather than use a color checker or grey card to empirically measure the ``true" white balance.~\cite{Liba2019}
+
+Additionally, Liba et al.
+developed a new error metric for training the model that better deals with heavily tinted illuminants which are common in night scenes (like those from colorful neon lights).
+The issue with traditional error metrics in low light is that they are based on how well the algorithm recovers \emph{all} of the color channels of the illuminant.
+This works well for the close to white true illumination of brighter scenes but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
+Such an image would look the same under all possible scalings of that channel.
+When a color channel is ``missing" like this it can produce inaccurately low accuracy results from the error metric.
+It is also unclear how to set the missing channels in the ground-truth illuminant data.~\cite{Liba2019}
+
+The existing error metric for color constancy looks at the error in appearance of a white patch in the image, but that idea doesn't work in heavily-tinted scenes with missing color channels so the improved error metric considers the appearance of an \emph{average} portion of the image under the recovered illumination.
+It is able to be less sensitive to errors in channels with lower mean values in the true image.~\cite{Liba2019}
+
+\begin{figure}
+\centering
+%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
+\begin{subfigure}{1.5in}
+\centering
+\includegraphics[width=1.5in]{figures/0GMA_20180814_233948_174_qc.jpg}
+\caption{Pixel default AWB}
+\label{fig:whitebalance:default}
+\end{subfigure}
+\begin{subfigure}{1.5in}
+\centering
+\includegraphics[width=1.5in]{figures/0GMA_20180814_233948_174_gcam.jpg}
+\caption{Liba et al.}
+\label{fig:whitebalance:liba}
+\end{subfigure}
+
+\caption{A comparison of the default implementation of FFCC in the Pixel's camera and the low-light optimized version by Liba et al.~\cite{blog:Levoy2018}}
+
+\label{fig:whitebalance}
+\end{figure}
 
@@
 
-Tone mapping is the process of mapping colors from a high-dynamic-range image to a medium with a more limited dynamic range \cite{wiki:ToneMapping}.
+\begin{figure}
+\centering
+\includegraphics[width=3in]{figures/Wright_of_Derby_The_Orrery}
+\caption{ \emph{Philosopher Lecturing on the Orrery}, by Joseph Wright of Derby, 1766 \cite{file:theOrrery}.
+The artist depicts a dark scene with bright, colorful detail while still maintaining the nighttime aesthetic by increasing contrast, surrounding the scene with darkness, and making the shadow areas completely black.~\cite{blog:Wronski2018}}
+\label{fig:theOrrery}
+\end{figure}
+
+Tone mapping is the process of mapping colors from a high-dynamic-range image to another in a medium with a more limited dynamic range \cite{wiki:ToneMapping}. 
 This is usually done by applying what are called tone mapping operators (TMOs).
 Some TMOs attempt to create results close to human vision, while others produce a more artistic rendition.
-Human vision loses color sensitivity as light levels are reduced.
-Liba et al.
-developed a TMO that maintains vibrant color in dark scenes without looking artificial with inspiration from Artistic techniques to maintain a nighttime aesthetic.
+While human vision loses color sensitivity and spatial acuity as light levels are reduced, Liba et al.
+intended to render vibrant and colorful images in low-light rather than emulating human vision.~\cite{Liba2019}
+
+Simply brightening the whole image (Figure~\ref{fig:toneMapping:clahe}) can result in low contrast and undesired saturated regions that make the image look flat.
+For centuries, artists have evoked a nighttime aesthetic through various methods such as the use of darker pigments, suppressed shadows, and increased contrast (Figure~\ref{fig:theOrrery}).
+Liba et al. developed a TMO, inspired by artistic techniques, that maintains vibrant color in dark scenes without looking artificial and while maintaining a nighttime aesthetic.
+
 Their TMO uses various heuristics on top of the tone mapping of Hasinoff et al.
 \cite{Hasinoff2016}.
-These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.
-\cite{Liba2019}
+These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.~\cite{Liba2019}
 
-\todo[inline]{explain these heuristics in more detail}
+%\todo[inline]{explain these heuristics in more detail}
 
@@
 
-\caption{An example of tone mapping \cite{Liba2019}}
-\todo[inline]{needs better caption}
+\caption{An example of tone mapping a nighttime scene.
+(a) The tone mapping of Hasinoff et al.
+\cite{Hasinoff2016} produces too dark of an image whereas using a different tone mapping technique (b) that brightens using histogram equalization (CLAHE~\cite{CLAHE})results in more detail but lacking global contrast.
+The tone mapping from Liba et al.
+(c) retains detail but still has the global contrast and dark areas to look like a night scene.~\cite{Liba2019}}
+
 % A nighttime scene (about 0.4 lux) that demonstrates the difficulty
@@
 
-\subsubsection{Limitations}
+\begin{figure*}
+\centering
+\begin{subfigure}{2.2in}
+\centering
+\includegraphics[width=2.2in]{figures/liba2019-figure-1a-95quality.jpg}
+\caption{Hasinoff et al.}
+\label{fig:libaResults:hasinoff}
+\end{subfigure}
+\begin{subfigure}{2.2in}
+\centering
+\includegraphics[width=2.2in]{figures/liba2019-figure-1b-95quality.jpg}
+\caption{Hasinoff et al. brightened}
+\label{fig:libaResults:Hasinoff}
+\end{subfigure}
+\begin{subfigure}{2.2in}
+\centering
+\includegraphics[width=2.2in]{figures/liba2019-figure-1c-95quality.jpg}
+\caption{Liba et al.}
+\label{fig:libaResults:liba}
+\end{subfigure}
+
+\caption{A comparison of images captured using the Hasinoff et al.
+pipeline (a and b) and the pipeline by Liba et al.
+showing the improvements in detail and noise from selecting a longer exposure time due to the low motion in the scene and taking more frames.
+It also shows the improvements in color reproduction from the improved white balance algorithm and tone mapping.~\cite{Liba2019}}
+
+\label{fig:libaResults}
+\end{figure*}
+
+This system was launched in November 2018 on Pixel phones as the "Night Sight" mode in the camera app.
+On Pixel, Pixel 2, and Pixel 3a it uses a burst merging technique adapted from Hasiniff et al.
+\cite{Hasinoff2016} to better handle motion while on faster Pixel phones it uses the super-resolution merging algorithm described by Wronski et al.
+\cite{Wronski2019} (Section~\ref{sec:handheldSuperRes}).
+The comparisons and results here all use the former merging technique.
+\cite{Liba2019}
+
+Liba et al.
+compared their system to the pipeline it built upon from Hasiniff et al.
+\cite{Hasinoff2016}.
+An example of their system producing lower noise, more detail, and more pleasing colors is shown in Figure~\ref{fig:libaResults}.
+They also compared their pipeline with neural network that operates on single raw images by Chen et al.~\cite{Chen2018} using frames from a similar camera to the one the neural network was trained on rather than a smartphone camera and still the system by Liba et al.
+produced higher-quality images.
+Additionally, Liba et al.'s system processes images much faster and with much less memory than the neural network by Chen et al.
+\cite{Chen2018}.~\cite{Liba2019}
+
+More recently, Night Sight has added an astrophotography mode allowing for even longer exposures (as long as 1-4 minutes) which allows for sharp and clear images of stars and extremely dark landscapes~\cite{blog:Kainz2019}.
+
+%\subsubsection{Limitations}
 
@@
 \section{Conclusions}
+
+New software techniques like these have allowed Google to continue using the same camera sensor hardware over multiple generations of their Pixel phones while continuing to be highly rated in photo quality.
+It has also enabled updates that continually improve the camera performance of previous Pixel phones to the point where they compete with the latest from other manufacturers.
 

### 2020-11-12 paper.tex
Updates from Overleaf

@@
 \graphicspath{ {./figures/} }
+
+\usepackage{amsmath}
+
+\usepackage{dblfloatfix}    % To enable figures at the bottom of page
 
@@
 Smartphone cameras present many challenges, most of which come from the need for them to be physically small.
-Their small size puts a fundamental limit on their ability to resolve detail and especially affects the ability to zoom and low-light photography.
+Their small size puts a fundamental limit on their ability to resolve detail and collect light, which makes low-light photography and zooming difficult.
 This paper presents two approaches to improve smartphone photography through software techniques.
-The first is handheld super-resolution which uses natural hand movement to improve the resolution of burst images.
+The first is handheld super-resolution which uses natural hand movement to improve the resolution smartphone images, especially when zoomed.
 The second approach is a system which improves low light photography in smartphones.
@@
 
+\begin{figure}
+\centering
+%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
+\includegraphics[width=20pc]{Bayer_pattern_on_sensor}
+\caption{A Bayer pattern on a sensor in isometric perspective.~\cite{wiki:BayerFilter}}
+\label{fig:BayerPattern}
+\end{figure}
+
+\begin{figure*}[t]
+\centering
+% TODO: Size correctly for width
+%\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
+\includegraphics[width=42pc]{Wronski2019-figure-2}
+\caption{An overview of the approach used by Wronski et al.~\cite{Wronski2019}.
+The initial burst of input frames (a) are aligned (d) to a base frame.
+The local features in the frames (b) are used to create kernels (c) (Section~\ref{sec:kernelReconstruction}) which are used along with the motion robustness model (f) (Section~\ref{sec:robustnessModel}) to combine (g) the frames separately for each color channel.
+The final image (h) is produced by normalizing the results of each channel.}
+\label{fig:Wronski2019Fig2}
+\end{figure*}
+
+\pagebreak
+
 \subsection{Burst photography}
@@
 A Bayer filter is a type of color filter array (CFA).
-The majority of digital image sensors in digital cameras and phones use a Bayer filter mosaic pattern to arrange RGB color filters on the sensor.
+CFAs are needed because the light sensing components in a digital image sensor can only detect the presence of light and not what specific color it is.
+The majority of digital image sensors in digital cameras and phones use a Bayer filter mosaic pattern to arrange RGB color filters on the sensor (Figure~\ref{fig:BayerPattern}).
 The pattern consists of 50\% green, 25\% red, and 25\% blue pixels.
@@
 Due to this pattern on the sensor, the raw output of a digital camera also has each pixel filtered to only red, green, or blue and a demosaicing algorithm must be used to interpolate the other values for each pixel.~\cite{wiki:BayerFilter}
-
-\begin{figure}
-\centering
-%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
-\includegraphics[width=3in]{Bayer_pattern_on_sensor}
-\caption{A Bayer pattern on a sensor in isometric perspective.~\cite{wiki:BayerFilter}}
-\label{fig:BayerPattern}
-\end{figure}
-
-\begin{figure*}[t]
-\centering
-% TODO: Size correctly for width
-%\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
-\includegraphics[width=7in]{Wronski2019-figure-2}
-\caption{An overview of the approach used by Wronski et al.~\cite{Wronski2019}.
-The initial burst of input frames (a) are aligned (d) to a base frame.
-The local features in the frames (b) are used to create kernels (c) (Section~\ref{sec:kernelReconstruction}) which are used along with the motion robustness model (f) (Section~\ref{sec:robustnessModel}) to combine (g) the frames separately for each color channel.
-The final image (h) is produced by normalizing the results of each channel.}
-\label{fig:Wronski2019Fig2}
-\end{figure*}
 
@@
 
-\todo[inline]{TODO: Explain kernels either here or in later section where it comes up}
+A kernel (or convolution matrix) in image processing refers to a small matrix that is used to apply effects like blurring or detect features in an image like edges.
+The kernel is applied by doing a convolution, adding each pixel of the image to its local neighbors, weighted by the kernel.
+\cite{wiki:kernel}
+
+For example, the following applies a kernel (the first matrix) to an image piece (the second matrix):
+\begin{equation*}
+\begin{bmatrix}
+a & b & c \\
+d & e & f \\
+g & h & i
+\end{bmatrix}
+*
+\begin{bmatrix}
+1 & 2 & 3 \\
+4 & 5 & 6 \\
+7 & 8 & 9
+\end{bmatrix}
+\end{equation*}
+
+The pixel in the center of the final image (coordinates $[2,2]$) would be this combination of the values of the image matrix weighted by the kernel matrix:
+\begin{equation*}
+(i \cdot 1)+(h \cdot 2)+(g \cdot 3)+(f \cdot 4)+(e \cdot 5)+(d \cdot 6)+(c \cdot 7)+(b \cdot 8)+(a \cdot 9)
+\end{equation*}
 
@@
 \subsection{Algorithm overview}
+
+\begin{figure*}
+\centering
+% TODO: Size correctly for width
+\includegraphics[width=42pc]{Wronski2019-figure-4}
+\caption{An illustration of subpixel displacements from a burst of four frames with linear hand motion.
+Each frame is offset by half a pixel on the x-axis and a quarter pixel on the y-axis from the last frame.
+After alignment, the pixel centers (black dots) uniformly cover the image with greater density than a single frame.~\cite{Wronski2019}}
+\label{fig:Wronski2019Fig4}
+
+% Illustration of a burst of four frames with linear hand motion. Each frame is offset from the
+% previous frame by half a pixel along the x-axis and a quarter pixel along the y-axis due to the hand motion. Afuer alignment to the base frame, the pixel
+% centers (black dots) uniformly cover the resampling grid (grey lines) at an increased density. In practice, the distribution is more random than in this simplified
+% example.
+
+\end{figure*}
 
@@
 
-\begin{figure*}
-\centering
-% TODO: Size correctly for width
-\includegraphics[width=7in]{Wronski2019-figure-4}
-\caption{An illustration of subpixel displacements from a burst of four frames with linear hand motion.
-Each frame is offset by half a pixel on the x-axis and a quarter pixel on the y-axis from the last frame.
-After alignment, the pixel centers (black dots) uniformly cover the image with greater density than a single frame.~\cite{Wronski2019}}
-\label{fig:Wronski2019Fig4}
-
-% Illustration of a burst of four frames with linear hand motion. Each frame is offset from the
-% previous frame by half a pixel along the x-axis and a quarter pixel along the y-axis due to the hand motion. Afuer alignment to the base frame, the pixel
-% centers (black dots) uniformly cover the resampling grid (grey lines) at an increased density. In practice, the distribution is more random than in this simplified
-% example.
-
-\end{figure*}
-
 One of the important conditions for multi-frame super-resolution is that the input contains multiple images that are sampled at different subpixel offsets.
@@
 Wronski et al.
-show how this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
+determined this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
 
-Wronski et al. analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography.
-%They used the rotational measurements from the phone's gyroscope but ignored the effect of translations in the analysis.
+Wronski et al. analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography using the rotational measurements from the phone's gyroscope.
 Their analysis showed that hand movement creates uniformly random angular displacements and relatively slow rotation of the device.
 
-While hand shake over a long time is uniformly random, over a short burst it could be more of a straight line.
-Wronski et al. show that this also provides a sufficiently uniform distribution of subpixel samples.
-Using the equidistribution theorem, which states that the sequence $\{a,2a,3a,\dotsc \bmod 1\}$ is uniformly distributed (assuming $a$ is an irrational number), with each pixel as a point sample and a least random scenario (the hand motion is regular and linear), the samples from all frames combined will be approximately uniformly distributed within the subpixel space (Figure~\ref{fig:Wronski2019Fig4}).
+%While hand shake over a long time is uniformly random, over a short burst it could be more of a straight line.
+%Wronski et al. show that this also provides a sufficiently uniform distribution of subpixel samples.
+%Using the equidistribution theorem, which states that the sequence $\{a,2a,3a,\dotsc \bmod 1\}$ is uniformly distributed (assuming $a$ is an irrational number), with each pixel as a point sample and a least random scenario (the hand motion is regular and linear), the samples from all frames combined will be approximately uniformly distributed within the subpixel space (Figure~\ref{fig:Wronski2019Fig4}).
 
-Wronski et al.
+%Wronski et al.
 also tested this concept empirically by measuring subpixel offsets by registration across 20 handheld bursts.
@@
 \subsection{Proposed super-resolution approach}
+
+\begin{figure}
+\centering
+\includegraphics[width=18pc]{figures/Wronski2019-figure-8.pdf}
+\caption{Plots of relative weights as a function of local features. The scale on the right indicates the weight given to each pixel in the $3\times 3$ neighborhood.~\cite{Wronski2019}}
+\label{fig:Wronski2019Fig8}
+% Merge kernels: Plots of relative weights in difgerent 3 × 3 sampling
+% kernels as a function of local tensor features
+\end{figure}
 
@@
 The process for each pixel in each color channel can be described as:
-\begin{equation}
+\begin{equation*}
 C(x, y)=\frac{\sum_{n} \sum_{i} c_{n, i} \cdot w_{n, i} \cdot \hat{R}_{n}}{\sum_{n} \sum_{i} w_{n, i} \cdot \hat{R}_{n}}
-\end{equation}
+\end{equation*}
 
@@
 %$\hat{R}_{n}$ is the local motion robustness score  described in Section~\ref{sec:robustnessModel}. \cite{Wronski2019}
-
-\begin{figure}
-\centering
-\includegraphics[width=3in]{figures/Wronski2019-figure-8.pdf}
-\caption{Plots of relative weights as a function of local features~\cite{Wronski2019}}
-\label{fig:Wronski2019Fig8}
-% Merge kernels: Plots of relative weights in difgerent 3 × 3 sampling
-% kernels as a function of local tensor features
-\end{figure}
 
@@
 \centering
-\includegraphics[width=3in]{figures/wronski2019-figure-9-95quality.jpg}
+\includegraphics[width=20pc]{figures/wronski2019-figure-9-95quality.jpg}
 \caption{A photograph of a moving bus: \textbf{Left}: Without a motion robustness model there are alignment errors and occlusions that result in tiling and ghosting artifacts.
@@
 To combine frames robustly, a confidence level is assigned to the local neighborhood of each pixel; the map of these confidences is called a \emph{robustness map}.
+
+This confidence level is assigned by computing the standard deviation in the image and a color difference between the base frame and the aligned input frame.
+Regions of this frame with differences smaller than the local standard deviation and those that are close to a pre-defined fraction of the spatial standard deviation will be merged while larger differences are likely non-aligned motion and are discarded.
+
+An example of what this can be seen in the middle image of figure~\ref{fig:Wronski2019Fig9}, the dark areas have a lower robustness value and the lighter areas have a greater robustness value.
+
 % try to reword below
-A robustness value of one corresponds to fully merged areas while a value of zero means rejected areas. The base frame is assigned a value of 1 because it is what the other frames are aligned to.
+%A robustness value of one corresponds to fully merged areas while a value of zero means rejected areas.
+An example of what this can be seen in the middle image of figure~\ref{fig:Wronski2019Fig9}, the dark areas have a lower robustness value and the lighter areas have a greater robustness value.
+The base frame is assigned a value of 1 because it is what the other frames are aligned to.
 
-The robustness values are determined by a statistical robustness model which must distinguish between aliasing, as it is needed for super-resolution, and frame misalignment, which is detrimental to super-resolution.
+%The robustness values are determined by a statistical robustness model which must distinguish between aliasing, as it is needed for super-resolution, and frame misalignment, which is detrimental to super-resolution.
 Wronski et al.
@@
 
-\todo[inline]{TODO: I need to explain specifics of how the motion robustness model works}
+%The statisitical robustness model computes the standard deviation in the images and a color difference between the base frame and the aligned input frame.
+Regions with smaller differences than the local standard deviation are merged.
 
@@
 
-\todo[inline]{TODO: Results and limitations}
+\begin{figure}[]
+\centering
+\includegraphics[width=20pc]{figures/wronski2019-figure-1-100quality-crop.jpg}
 
-\subsubsection{Limitations}
+\caption{A comparison of the traditional frame merging approach (\textbf{top}) and the approach by Wronski et al.
+(\textbf{bottom}).
+The super resolution merge algorithm improves sharpness and detail while eliminating the
+Moiré aliasing artifacts seen in the third (red) inset.~\cite{Wronski2019}}
+
+\label{fig:wronskiResults}
+\end{figure}
+
+Wronski et al.
+used a variety of methods to evaluate their algorithm and compare it with other demosaicing, merging, and super-resolution techniques.
+They also compared their method as part of a full camera pipeline to that of Hasinoff et al.
+\cite{Hasinoff2016}.
+Figure~\ref{fig:wronskiResults} shows one of these comparison between a traditional method that merges frames of each color channel first and then applies demosaicing to the method developed by Wronski et al.
+\cite{Wronski2019}
 
@@
 
-When capturing a burst of images on a smartphone, the exposure time and sensor gain (ISO) needs to be selected for each frame.
-Liba et al.
-use the same strategy as in~\cite{Hasinoff2016} and capture all frames in the burst with the same exposure time and ISO.
-For effective low-light photography these settings need to be automatically selected within the constraints of keeping the total capture time low ($\leq$ 6 seconds) and the total number of frames within the device's memory limits.
-
-This process consists of splitting the target sensitivity, which is based on the brightness of the scene, into exposure time and gain and calculating the number of frames with respect to the time and memory limits.
-The traditional method used in Hasinoff et al.~\cite{Hasinoff2016} uses a fixed ``exposure schedule" that simply keeps the exposure time low to limit motion blur.
-This method usually works well but can be improved for low-light photography.
-
-The ``motion metering" described by Liba et al.
-selects the exposure time and gain by predicting future motion in the scene and motion of the camera itself.
-This is used to select slower exposures for scenes with no motion and shorter exposures for those with motion to reduce motion blur when needed and increase signal-to-noise ration when possible.
-
-%\subsubsection{Motion prediction}
-
-\subsubsection{Stability detection}
-
-More experienced photographers often brace their device against a surface or put it on a tripod in low-light situations.
-If this can be detected, longer exposures can be used without risk of motion blur.
-Liba et al.'s motion metering system captures up to 333 ms exposures when handheld but when stabilized it captures up to 1 second exposures.
-To detect a stabilized device, measurements from the device's gyroscope are averaged over a short time, 1.466 s, before the shutter button was pressed while masking out the last 0.4 s of the measurements to prevent vibrations from the user pressing the shutter button from affecting the result.
-
-%\subsubsection{Exposure time selection}
-
 \begin{figure}
 \centering
-\includegraphics[width=3in]{figures/liba2019-figure-7.pdf}
+\includegraphics[width=20pc]{figures/liba2019-figure-7.pdf}
 \caption{A traditional static exposure schedule compared to the dynamic exposure scheudle of Liba et al. at various levels of motion.~\cite{Liba2019}}
@@
 
+When capturing a burst of images on a smartphone, the exposure time and sensor gain (the sensitivity of the sensor to light, also known as ISO) needs to be selected for each frame.
+Liba et al.
+use the same strategy as in~\cite{Hasinoff2016} and capture all frames in the burst with the same exposure time and ISO.
+For effective low-light photography these settings need to be automatically selected within the constraints of keeping the total capture time low ($\leq$ 6 seconds) and the total number of frames within the device's memory limits.
+
+This process consists of splitting the target sensitivity, which is based on the brightness of the scene, into exposure time and gain and calculating the number of frames with respect to the time and memory limits.
+The traditional method used in Hasinoff et al.~\cite{Hasinoff2016} uses a fixed ``exposure schedule" that simply keeps the exposure time low to limit motion blur.
+This method usually works well but can be improved for low-light photography.
+
+The ``motion metering" described by Liba et al.
+selects the exposure time and gain by predicting future motion in the scene and motion of the camera itself.
+It produces a variable exposure schedule that varies based on the amount of motion detected (Figure~\ref{fig:exposure}).
+This is used to select slower exposures for scenes with no motion and shorter exposures for those with motion to reduce motion blur when needed and increase signal-to-noise ratio when possible.
+
+%\subsubsection{Motion prediction}
+
+\subsubsection{Stability detection}
+
+More experienced photographers often brace their device against a surface or put it on a tripod in low-light situations.
+If this can be detected, longer exposures can be used without risk of motion blur.
+Liba et al.'s motion metering system captures up to 333 ms exposures when handheld but when stabilized it captures up to 1 second exposures.
+To detect a stabilized device, measurements from the device's gyroscope are averaged over a short time, 1.466 s, before the shutter button was pressed while masking out the last 0.4 s of the measurements to prevent vibrations from the user pressing the shutter button from affecting the result.
+
+%\subsubsection{Exposure time selection}
+
 %\subsection{Motion-adaptive burst merging}
@@
 
-Humans perceive color correctly of things even when lit with colored light, a phenomenon called color constancy.
-This perception can break down when a photograph is taken under one type of light and viewed under different light.
-It will look tinted (Figure~\ref{fig:whitebalance:default}).
-Cameras correct for this by determining the color of the majority of the illumination in the scene and correcting the colors in the image such that they appear to be lit by a neutral (white) illumination.
-This Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image. \cite{Liba2019, blog:Levoy2018}
-
-The current best-performing color constancy algorithm is ``Fast Fourier Color Constancy" (FFCC) \cite{Barron2017}.
-It is a machine learning based algorithm that ``works by learning and then applying a filter to a histogram of log-chrominance pixel and edge intensities"~\cite{Liba2019}.
-Liba et al.
-trained the FFCC algorithm with a new dataset and error metric to better handle challenging low-light scenes (Figure~\ref{fig:whitebalance}).
-
-Liba et al.
-collected 5000 images using mobile devices of various scenes that demonstrate a range of light levels to be able to train their implementation of FFCC.
-To obtain better ground truth coloring for the training data they had real professional photographers manually white balance the images with the most ``aesthetically preferable" white balance for the scene rather than use a color checker or grey card to empirically measure the ``true" white balance.~\cite{Liba2019}
-
-Additionally, Liba et al.
-developed a new error metric for training the model that better deals with heavily tinted illuminants which are common in night scenes (like those from colorful neon lights).
-The issue with traditional error metrics in low light is that they are based on how well the algorithm recovers \emph{all} of the color channels of the illuminant.
-This works well for the close to white true illumination of brighter scenes but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
-Such an image would look the same under all possible scalings of that channel.
-When a color channel is ``missing" like this it can produce inaccurately low accuracy results from the error metric.
-It is also unclear how to set the missing channels in the ground-truth illuminant data.~\cite{Liba2019}
-
-The existing error metric for color constancy looks at the error in appearance of a white patch in the image, but that idea doesn't work in heavily-tinted scenes with missing color channels so the improved error metric considers the appearance of an \emph{average} portion of the image under the recovered illumination.
-It is able to be less sensitive to errors in channels with lower mean values in the true image.~\cite{Liba2019}
-
 \begin{figure}
 \centering
-%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
-\begin{subfigure}{1.5in}
+\begin{subfigure}{9pc}
 \centering
-\includegraphics[width=1.5in]{figures/0GMA_20180814_233948_174_qc.jpg}
+\includegraphics[width=9pc]{figures/0GMA_20180814_233948_174_qc.jpg}
 \caption{Pixel default AWB}
@@
 \end{subfigure}
-\begin{subfigure}{1.5in}
+\begin{subfigure}{9pc}
 \centering
-\includegraphics[width=1.5in]{figures/0GMA_20180814_233948_174_gcam.jpg}
+\includegraphics[width=9pc]{figures/0GMA_20180814_233948_174_gcam.jpg}
 \caption{Liba et al.}
@@
 
-\subsection{Tone mapping}
-\label{sec:toneMapping}
-
 \begin{figure}
 \centering
-\includegraphics[width=3in]{figures/Wright_of_Derby_The_Orrery}
+\includegraphics[width=20pc]{figures/Wright_of_Derby_The_Orrery}
 \caption{ \emph{Philosopher Lecturing on the Orrery}, by Joseph Wright of Derby, 1766 \cite{file:theOrrery}.
@@
 
-Tone mapping is the process of mapping colors from a high-dynamic-range image to another in a medium with a more limited dynamic range \cite{wiki:ToneMapping}. 
-This is usually done by applying what are called tone mapping operators (TMOs).
-Some TMOs attempt to create results close to human vision, while others produce a more artistic rendition.
-While human vision loses color sensitivity and spatial acuity as light levels are reduced, Liba et al.
-intended to render vibrant and colorful images in low-light rather than emulating human vision.~\cite{Liba2019}
+Humans perceive color correctly of things even when lit with colored light, a phenomenon called color constancy.
+This perception can break down when a photograph is taken under one type of light and viewed under different light.
+It will look tinted (Figure~\ref{fig:whitebalance:default}).
+Cameras correct for this by determining the color of the majority of the illumination in the scene and correcting the colors in the image such that they appear to be lit by a neutral (white) illumination.
+This Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image. \cite{Liba2019, blog:Levoy2018}
 
-Simply brightening the whole image (Figure~\ref{fig:toneMapping:clahe}) can result in low contrast and undesired saturated regions that make the image look flat.
-For centuries, artists have evoked a nighttime aesthetic through various methods such as the use of darker pigments, suppressed shadows, and increased contrast (Figure~\ref{fig:theOrrery}).
-Liba et al. developed a TMO, inspired by artistic techniques, that maintains vibrant color in dark scenes without looking artificial and while maintaining a nighttime aesthetic.
+The current best-performing color constancy algorithm is the machine learning based ``Fast Fourier Color Constancy" (FFCC) \cite{Barron2017}.
+%It is a machine learning based algorithm that ``works by learning and then applying a filter to a histogram of log-chrominance pixel and edge intensities"~\cite{Liba2019}. 
+Liba et al. trained the FFCC algorithm with a new dataset and error metric to better handle challenging low-light scenes (Figure~\ref{fig:whitebalance}).
 
-Their TMO uses various heuristics on top of the tone mapping of Hasinoff et al.
-\cite{Hasinoff2016}.
-These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.~\cite{Liba2019}
-
-%\todo[inline]{explain these heuristics in more detail}
+Liba et al.
+collected 5000 images using mobile devices of various scenes that demonstrate a range of light levels to be able to train their implementation of FFCC.
+To obtain better ground truth coloring for the training data they had real professional photographers manually white balance the images with the most ``aesthetically preferable" white balance for the scene rather than use a color checker or grey card to empirically measure the ``true" white balance.~\cite{Liba2019}
 
@@
 \centering
-%\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
-\begin{subfigure}{1in}
+\begin{subfigure}{6pc}
 \centering
-\includegraphics[width=1in]{figures/liba2019-figure-14a-95quality.jpg}
+\includegraphics[width=6pc]{figures/liba2019-figure-14a-95quality.jpg}
 \caption{Baseline}
@@
 \end{subfigure}
-\begin{subfigure}{1in}
+\begin{subfigure}{6pc}
 \centering
-\includegraphics[width=1in]{figures/liba2019-figure-14b-95quality.jpg}
+\includegraphics[width=6pc]{figures/liba2019-figure-14b-95quality.jpg}
 \caption{CLAHE}
@@
 \end{subfigure}
-\begin{subfigure}{1in}
+\begin{subfigure}{6pc}
 \centering
-\includegraphics[width=1in]{figures/liba2019-figure-14c-95quality.jpg}
+\includegraphics[width=6pc]{figures/liba2019-figure-14c-95quality.jpg}
 \caption{Liba et al.}
@@
 
-\subsection{Results}
-
 \begin{figure*}
 \centering
-\begin{subfigure}{2.2in}
+\begin{subfigure}{13pc}
 \centering
-\includegraphics[width=2.2in]{figures/liba2019-figure-1a-95quality.jpg}
+\includegraphics[width=13pc]{figures/liba2019-figure-1a-95quality.jpg}
 \caption{Hasinoff et al.}
@@
 \end{subfigure}
-\begin{subfigure}{2.2in}
+\begin{subfigure}{13pc}
 \centering
-\includegraphics[width=2.2in]{figures/liba2019-figure-1b-95quality.jpg}
+\includegraphics[width=13pc]{figures/liba2019-figure-1b-95quality.jpg}
 \caption{Hasinoff et al. brightened}
@@
 \end{subfigure}
-\begin{subfigure}{2.2in}
+\begin{subfigure}{13pc}
 \centering
-\includegraphics[width=2.2in]{figures/liba2019-figure-1c-95quality.jpg}
+\includegraphics[width=13pc]{figures/liba2019-figure-1c-95quality.jpg}
 \caption{Liba et al.}
@@
 \end{figure*}
+
+Additionally, Liba et al.
+developed a new error metric for training the model that better deals with heavily tinted illuminants which are common in night scenes (like those from colorful neon lights).
+The issue with traditional error metrics in low light is that they are based on how well the algorithm recovers \emph{all} of the color channels of the illuminant.
+This works well for the close to white true illumination of brighter scenes but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
+Such an image would look the same under all possible transformations of that channel.
+When a color channel is ``missing" like this it can produce inaccurately low accuracy results from the error metric.
+It is also unclear how to set the missing channels in the ground-truth illuminant data.~\cite{Liba2019}
+
+The existing error metric for color constancy looks at the error in appearance of a white patch in the image, but that idea doesn't work in heavily-tinted scenes with missing color channels so the improved error metric considers the appearance of an \emph{average} portion of the image under the recovered illumination.
+It is able to be less sensitive to errors in channels with lower mean values in the true image.~\cite{Liba2019}
+
+\subsection{Tone mapping}
+\label{sec:toneMapping}
+
+Tone mapping is the process of mapping colors from a high-dynamic-range image to another in a medium with a more limited dynamic range \cite{wiki:ToneMapping}. 
+This is usually done by applying what are called tone mapping operators (TMOs).
+Some TMOs attempt to create results close to human vision, while others produce a more artistic rendition.
+While human vision loses color sensitivity and spatial acuity as light levels are reduced, Liba et al.
+still wanted to render vibrant and colorful images in low-light rather than emulating human vision.~\cite{Liba2019}
+
+Simply brightening the whole image (Figure~\ref{fig:toneMapping:clahe}) can result in low contrast and undesired saturated regions that make the image look flat.
+For centuries, artists have evoked a nighttime aesthetic through various methods such as the use of darker pigments, suppressed shadows, and increased contrast (Figure~\ref{fig:theOrrery}).
+Liba et al. developed a TMO, inspired by artistic techniques, that maintains vibrant color in dark scenes without looking artificial and while maintaining a nighttime aesthetic.
+
+Their TMO uses various heuristics on top of the tone mapping of Hasinoff et al.
+\cite{Hasinoff2016}.
+These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.~\cite{Liba2019}
+
+%\todo[inline]{explain these heuristics in more detail}
+
+\subsection{Results}
 
@@
 
-%\subsubsection{Limitations}
-
 % ----------------------------------------
@@
 It has also enabled updates that continually improve the camera performance of previous Pixel phones to the point where they compete with the latest from other manufacturers.
+Computational photography will continue to become an ever more important part of cameras, both in smartphones and in other applications.
 
 \section*{Acknowledgments}
+
+I would like to thank Nic McPhee, Elena Machkasova, and Kristin Shirley for their feedback and advice.
+
 \label{sec:acknowledgments}
+
+\pagebreak
 

### 2020-11-15 paper.tex
Updates from Overleaf

@@
 \usepackage{dblfloatfix}    % To enable figures at the bottom of page
+
+\usepackage[
+bookmarksnumbered,
+hidelinks,
+pdflang={en},
+pdfdisplaydoctitle,
+pdfauthor={Paul Friederichsen},
+pdftitle={Recent Advances in Smartphone Computational Photography},
+pdfkeywords={computational photography, super-resolution, image processing, low-light imaging}
+]{hyperref}
+
+%\urlstyle{rm}
+
+\boilerplate={This work is licensed under the Creative Commons 
+Attribution-NonCommercial-ShareAlike 4.0 International License. 
+To view a copy of this license, visit 
+\href{https://creativecommons.org/licenses/by-nc-sa/4.0/}{creativecommons.org/licenses/by-nc-sa/4.0/}.}
 
@@
 	\affaddr{Morris, Minnesota, USA 56267}\\
-	\email{fried701@morris.umn.edu}
+	\email{\href{mailto:fried701@morris.umn.edu}{fried701@morris.umn.edu}
+	}
 }
@@
 This means their ability to gather light is significantly reduced compared to larger dedicated cameras.
-Most modern smartphones use a burst system to capture multiple images when the user presses the capture button and merges them to improve image quality.
+Most modern smartphones use a burst system to capture multiple images and merge them together to improve image quality.
 
@@
 %\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
-\includegraphics[width=20pc]{Bayer_pattern_on_sensor}
+\includegraphics[width=\columnwidth]{Bayer_pattern_on_sensor}
 \caption{A Bayer pattern on a sensor in isometric perspective.~\cite{wiki:BayerFilter}}
@@
 %\psfig{file=figures/Wronski2019-figure-2.pdf,width =7in}
-\includegraphics[width=42pc]{Wronski2019-figure-2}
+\includegraphics[width=\textwidth]{Wronski2019-figure-2}
 \caption{An overview of the approach used by Wronski et al.~\cite{Wronski2019}.
@@
 An image processing pipeline refers to the process the raw data from a camera sensor goes through to be turned into the final image file that can be displayed and shared. 
-
 Most smartphones use a burst processing pipeline for their cameras.
-Generally, burst processing involves taking a series of raw exposures and merging them together to form the final image.
+Generally, burst processing involves taking a series of exposures and merging them together to form the final image.
 Most smartphones operate in a \emph{zero-shutter lag} mode by default.
-In this mode raw frames (the full unprocessed sensor output) are continuously captured to a temporary memory while the camera app is open.
-When the user presses the shutter button, several of the most recent frames are sent to the camera processing pipeline.
+In this mode, raw frames (the full unprocessed sensor output) are continuously captured to a temporary area in memory while the camera app is open.
+When the user presses the shutter button, several of the most recent frames are sent to the camera processing pipeline to be merged.
 
@@
 \cite{Hasinoff2016} which used bursts of constant low-exposure frames to increase dynamic range and signal-to-noise ratio.
+It specifically uses under-exposed frames to reduce motion blur.
 
@@
 % TODO: Size correctly for width
-\includegraphics[width=42pc]{Wronski2019-figure-4}
+\includegraphics[width=\textwidth]{Wronski2019-figure-4}
 \caption{An illustration of subpixel displacements from a burst of four frames with linear hand motion.
@@
 
-\subsection{Proposed super-resolution approach}
+%\subsection{Proposed super-resolution approach}
+
+%Super-resolution involves recreating a higher resolution signal from many lower resolution samples.
+Due to the random distribution of pixel shifts that result from hand movement, kernel regression was deemed a good technique for this reconstruction.
+Wronski et al.
+used anisotropic Gaussian Radial Basis Function (RBF) kernels (Section~\ref{sec:kernelReconstruction}) and a statistical robustness model (Section~\ref{sec:robustnessModel}) to combine burst frames into a higher resolution image.
+
+\subsection{Kernel reconstruction}
+\label{sec:kernelReconstruction}
 
@@
 \centering
-\includegraphics[width=18pc]{figures/Wronski2019-figure-8.pdf}
+\includegraphics[width=19pc]{figures/Wronski2019-figure-8.pdf}
 \caption{Plots of relative weights as a function of local features. The scale on the right indicates the weight given to each pixel in the $3\times 3$ neighborhood.~\cite{Wronski2019}}
@@
 
-Super-resolution involves recreating a higher resolution signal from many lower resolution samples.
-Due to the random distribution of pixel shifts that result from hand movement, kernel regression was deemed a good technique for this reconstruction.
-Wronski et al.
-used anisotropic Gaussian Radial Basis Function (RBF) kernels (Section~\ref{sec:kernelReconstruction}) and a statistical robustness model (Section~\ref{sec:robustnessModel}) to combine burst frames into a higher resolution image.
-
-\subsubsection{Kernel reconstruction}
-\label{sec:kernelReconstruction}
-
-The core idea of the algorithm is considering the pixels of multiple raw frames as randomly offset, aliased, and noisy measurements of three original continuous signals, one for each color channel.
+The core idea of the algorithm is considering the pixels of multiple raw frames with hand motion as randomly offset, aliased, and noisy measurements of three original continuous signals, one for each color channel.
 The algorithm creates the final output image pixel-by-pixel.
@@
 
-\subsubsection{Motion Robustness}
+\subsection{Motion robustness}
 \label{sec:robustnessModel}
@@
 \centering
-\includegraphics[width=20pc]{figures/wronski2019-figure-9-95quality.jpg}
+\includegraphics[width=\columnwidth]{figures/wronski2019-figure-9-95quality.jpg}
 \caption{A photograph of a moving bus: \textbf{Left}: Without a motion robustness model there are alignment errors and occlusions that result in tiling and ghosting artifacts.
@@
 
-\begin{figure}[]
-\centering
-\includegraphics[width=20pc]{figures/wronski2019-figure-1-100quality-crop.jpg}
-
-\caption{A comparison of the traditional frame merging approach (\textbf{top}) and the approach by Wronski et al.
-(\textbf{bottom}).
-The super resolution merge algorithm improves sharpness and detail while eliminating the
-Moiré aliasing artifacts seen in the third (red) inset.~\cite{Wronski2019}}
-
-\label{fig:wronskiResults}
-\end{figure}
-
 Wronski et al.
@@
 \cite{Wronski2019}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/wronski2019-figure-1-100quality-crop.jpg}
+
+\caption{A comparison of the traditional frame merging approach (\textbf{top}) and the approach by Wronski et al.
+(\textbf{bottom}).
+The super resolution merge algorithm improves sharpness and detail while eliminating the
+Moiré aliasing artifacts seen in the third (\textbf{red}) frame.~\cite{Wronski2019}}
+
+\label{fig:wronskiResults}
+\end{figure}
 
@@
 
-\subsection{Motion Metering}
+\subsection{Motion metering}
 \label{sec:motionMetering}
@@
 \centering
-\includegraphics[width=20pc]{figures/liba2019-figure-7.pdf}
-\caption{A traditional static exposure schedule compared to the dynamic exposure scheudle of Liba et al. at various levels of motion.~\cite{Liba2019}}
+\includegraphics[width=\columnwidth]{figures/liba2019-figure-7.pdf}
+\caption{A traditional static exposure schedule compared to the dynamic exposure schedule of Liba et al. at various levels of motion.~\cite{Liba2019}}
 \label{fig:exposure}
@@
 
+Additionally, experienced photographers often brace their device against a surface or put it on a tripod in low light.
+Liba et al.'s motion metering system is able to use even longer exposures (up to 1 second) by detecting this using measurements from the device's gyroscope.
+
 %\subsubsection{Motion prediction}
 
-\subsubsection{Stability detection}
+%\subsubsection{Stability detection}
 
-More experienced photographers often brace their device against a surface or put it on a tripod in low-light situations.
+%More experienced photographers often brace their device against a surface or put it on a tripod in low-light situations.
 If this can be detected, longer exposures can be used without risk of motion blur.
@@
 This is used on some devices while those with a fast enough system-on-chip use the merging method from Wronski et al.
-(Section~\ref{sec:handheldSuperRes}).
+(Section~\ref{setc:handheldSuperRes}).
 
@@
 
-\subsection{Auto white balance in low-light}
+\subsection{Auto white balance in low light}
 \label{sec:autoWhiteBal}
@@
 \centering
-\begin{subfigure}{9pc}
+\begin{subfigure}{9.8pc}
 \centering
-\includegraphics[width=9pc]{figures/0GMA_20180814_233948_174_qc.jpg}
+\includegraphics[width=9.8pc]{figures/0GMA_20180814_233948_174_qc.jpg}
 \caption{Pixel default AWB}
@@
 \end{subfigure}
-\begin{subfigure}{9pc}
+\begin{subfigure}{9.8pc}
 \centering
-\includegraphics[width=9pc]{figures/0GMA_20180814_233948_174_gcam.jpg}
+\includegraphics[width=9.8pc]{figures/0GMA_20180814_233948_174_gcam.jpg}
 \caption{Liba et al.}
@@
 \label{fig:whitebalance}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=20pc]{figures/Wright_of_Derby_The_Orrery}
-\caption{ \emph{Philosopher Lecturing on the Orrery}, by Joseph Wright of Derby, 1766 \cite{file:theOrrery}.
-The artist depicts a dark scene with bright, colorful detail while still maintaining the nighttime aesthetic by increasing contrast, surrounding the scene with darkness, and making the shadow areas completely black.~\cite{blog:Wronski2018}}
-\label{fig:theOrrery}
 \end{figure}
@@
 
+Additionally, Liba et al.
+developed a new error metric for training the model that better deals with heavily tinted illuminants which are common in night scenes (like those from colorful neon lights).
+The issue with traditional error metrics in low light is that they are based on how well the algorithm recovers \emph{all} of the color channels of the illuminant.
+This works well for the close to white true illumination of brighter scenes but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
+Such an image would look the same under all possible transformations of that channel.
+\pagebreak 
+When a color channel is ``missing" like this it can produce inaccurately low accuracy results from the error metric.
+It is also unclear how to set the missing channels in the ground-truth illuminant data.~\cite{Liba2019}
+
+The existing error metric for color constancy looks at the error in appearance of a white patch in the image, but that idea doesn't work in heavily-tinted scenes with missing color channels so the improved error metric considers the appearance of an \emph{average} portion of the image under the recovered illumination.
+It is able to be less sensitive to errors in channels with lower mean values in the true image.~\cite{Liba2019}
+
 \begin{figure}
 \centering
-\begin{subfigure}{6pc}
+\includegraphics[trim={0 000 0 0}, clip=true, width=\columnwidth]{figures/Wright_of_Derby_The_Orrery}
+\caption{ \emph{Philosopher Lecturing on the Orrery}, by Joseph Wright of Derby, 1766 \cite{file:theOrrery}.
+The artist depicts a dark scene with bright, colorful detail while still maintaining the nighttime aesthetic by increasing contrast, surrounding the scene in darkness, and keeping the shadow areas completely black.~\cite{blog:Wronski2018}}
+\label{fig:theOrrery}
+\end{figure}
+
+\begin{figure}
 \centering
-\includegraphics[width=6pc]{figures/liba2019-figure-14a-95quality.jpg}
+\begin{subfigure}{6.5pc}
+\centering
+\includegraphics[width=6.5pc]{figures/liba2019-figure-14a-95quality.jpg}
 \caption{Baseline}
@@
 \end{subfigure}
-\begin{subfigure}{6pc}
+\begin{subfigure}{6.5pc}
 \centering
-\includegraphics[width=6pc]{figures/liba2019-figure-14b-95quality.jpg}
+\includegraphics[width=6.5pc]{figures/liba2019-figure-14b-95quality.jpg}
 \caption{CLAHE}
@@
 \end{subfigure}
-\begin{subfigure}{6pc}
+\begin{subfigure}{6.5pc}
 \centering
-\includegraphics[width=6pc]{figures/liba2019-figure-14c-95quality.jpg}
+\includegraphics[width=6.5pc]{figures/liba2019-figure-14c-95quality.jpg}
 \caption{Liba et al.}
@@
 
-\caption{An example of tone mapping a nighttime scene.
-(a) The tone mapping of Hasinoff et al.
-\cite{Hasinoff2016} produces too dark of an image whereas using a different tone mapping technique (b) that brightens using histogram equalization (CLAHE~\cite{CLAHE})results in more detail but lacking global contrast.
+\caption{An example of tone mapping a night scene.
+The tone mapping of Hasinoff et al.
+\cite{Hasinoff2016} (\textbf{a}) produces too dark of an image while using a different tone mapping technique (\textbf{b}) that brightens using histogram equalization (CLAHE~\cite{CLAHE}) results in more detail but lacking global contrast.
 The tone mapping from Liba et al.
-(c) retains detail but still has the global contrast and dark areas to look like a night scene.~\cite{Liba2019}}
+(\textbf{c}) retains detail while keeping global contrast and dark areas to look like a night scene.~\cite{Liba2019}}
 
@@
 \centering
-\begin{subfigure}{13pc}
+\begin{subfigure}{13.8pc}
 \centering
-\includegraphics[width=13pc]{figures/liba2019-figure-1a-95quality.jpg}
+\includegraphics[width=13.8pc]{figures/liba2019-figure-1a-95quality.jpg}
 \caption{Hasinoff et al.}
@@
 \end{subfigure}
-\begin{subfigure}{13pc}
+\begin{subfigure}{13.8pc}
 \centering
-\includegraphics[width=13pc]{figures/liba2019-figure-1b-95quality.jpg}
+\includegraphics[width=13.8pc]{figures/liba2019-figure-1b-95quality.jpg}
 \caption{Hasinoff et al. brightened}
@@
 \end{subfigure}
-\begin{subfigure}{13pc}
+\begin{subfigure}{13.8pc}
 \centering
-\includegraphics[width=13pc]{figures/liba2019-figure-1c-95quality.jpg}
+\includegraphics[width=13.8pc]{figures/liba2019-figure-1c-95quality.jpg}
 \caption{Liba et al.}
@@
 \end{figure*}
-
-Additionally, Liba et al.
-developed a new error metric for training the model that better deals with heavily tinted illuminants which are common in night scenes (like those from colorful neon lights).
-The issue with traditional error metrics in low light is that they are based on how well the algorithm recovers \emph{all} of the color channels of the illuminant.
-This works well for the close to white true illumination of brighter scenes but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
-Such an image would look the same under all possible transformations of that channel.
-When a color channel is ``missing" like this it can produce inaccurately low accuracy results from the error metric.
-It is also unclear how to set the missing channels in the ground-truth illuminant data.~\cite{Liba2019}
-
-The existing error metric for color constancy looks at the error in appearance of a white patch in the image, but that idea doesn't work in heavily-tinted scenes with missing color channels so the improved error metric considers the appearance of an \emph{average} portion of the image under the recovered illumination.
-It is able to be less sensitive to errors in channels with lower mean values in the true image.~\cite{Liba2019}
 
@@
 \cite{Hasinoff2016}.
-These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.~\cite{Liba2019}
+These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.
+It produces a more detailed image while maintaining global contrast and the nighttime aesthetic (Figure~\ref{fig:toneMapping:liba}).~\cite{Liba2019}
 
@@
 
-This system was launched in November 2018 on Pixel phones as the "Night Sight" mode in the camera app.
+This system was launched in November 2018 on Pixel phones as the ``Night Sight" mode in the camera app.
 On Pixel, Pixel 2, and Pixel 3a it uses a burst merging technique adapted from Hasiniff et al.
@@
 They also compared their pipeline with neural network that operates on single raw images by Chen et al.~\cite{Chen2018} using frames from a similar camera to the one the neural network was trained on rather than a smartphone camera and still the system by Liba et al.
-produced higher-quality images.
-Additionally, Liba et al.'s system processes images much faster and with much less memory than the neural network by Chen et al.
-\cite{Chen2018}.~\cite{Liba2019}
+produced higher-quality images and processed images much faster and using less memory.~\cite{Liba2019}
 
-More recently, Night Sight has added an astrophotography mode allowing for even longer exposures (as long as 1-4 minutes) which allows for sharp and clear images of stars and extremely dark landscapes~\cite{blog:Kainz2019}.
+More recently, an astrophotography mode has been added to Night Sight allowing for even longer exposures (as long as 1-4 minutes) which allows for sharp and clear images of stars and extremely dark landscapes~\cite{blog:Kainz2019}.
 
@@
 
-New software techniques like these have allowed Google to continue using the same camera sensor hardware over multiple generations of their Pixel phones while continuing to be highly rated in photo quality.
+These new computational photography techniques and others like them have allowed Google to continue using the same camera sensor hardware over multiple generations of their Pixel phones while continuing to be highly rated in photo quality.
 It has also enabled updates that continually improve the camera performance of previous Pixel phones to the point where they compete with the latest from other manufacturers.
-Computational photography will continue to become an ever more important part of cameras, both in smartphones and in other applications.
+Computational photography techniques will continue to be an ever more important part off cameras and image processing in smartphones, professional photography, and other applications.
 
@@
 \label{sec:acknowledgments}
-
-\pagebreak
 

### 2020-11-22 paper.tex
Updates from Overleaf

@@
 pdftitle={Recent Advances in Smartphone Computational Photography},
-pdfkeywords={computational photography, super-resolution, image processing, low-light imaging}
+pdfkeywords={computational photography, image processing, low-light imaging, photography, super-resolution}
 ]{hyperref}
@@
 
-\keywords{computational photography, super-resolution, image processing, low-light imaging}
+\keywords{computational photography, image processing, low-light imaging, photography, super-resolution}
 
@@
 In this section I explain several key concepts that are important for the work in Sections \ref{sec:handheldSuperRes} and \ref{sec:handheldLowLight}.
-This includes Burst photography (Section~\ref{sec:background:burstPhotography}), Bayer filters (Section~\ref{sec:background:bayerFilter}), demosaicing (Section~\ref{sec:background:demosaicing}), aliasing (Section~\ref{sec:background:aliasing}), super-resolution (Section~\ref{sec:background:superResolution}), and kernels (Section~\ref{sec:background:kernels}).
+This includes Burst photography (Section~\ref{sec:background:burstPhotography}), Bayer filters (Section~\ref{sec:background:bayerFilter}), demosaicing and aliasing (Section~\ref{sec:background:demosaicing}), super-resolution (Section~\ref{sec:background:superResolution}), and kernels (Section~\ref{sec:background:kernels}).
 
@@
 %\psfig{file=figures/Bayer_pattern_on_sensor.pdf,width =3in}
-\includegraphics[width=\columnwidth]{Bayer_pattern_on_sensor}
+\includegraphics[width=\columnwidth]{20190404183627!Bayer_pattern_on_sensor}
 \caption{A Bayer pattern on a sensor in isometric perspective.~\cite{wiki:BayerFilter}}
@@
 
-\subsection{Burst photography}
+\subsection{Burst Photography}
 \label{sec:background:burstPhotography}
@@
 
-\subsection{Bayer filter}
+\subsection{Bayer Filters}
 \label{sec:background:bayerFilter}
@@
 
-\subsection{Demosaicing}
+\subsection{Demosaicing and Aliasing}
 \label{sec:background:demosaicing}
@@
 This process means two-thirds of the final image is reconstructed from the available data.
-The demosaicing process may introduce various artifacts in the final image due to aliasing.~\cite{blog:Wronski2018}
+The demosaicing process may introduce various artifacts in the final image due to aliasing, an effect that happens when the camera sensor is unable to correctly represent the patterns and details present in a scene due to its resolution.
+One type of artifact this causes are Moiré patterns, a type of interference pattern.~\cite{blog:Wronski2018}
 %These typically include false color artifacts like zippering and Moiré patterns~\cite{Wronski2019}.
@@
 
-\subsection{Aliasing}
-\label{sec:background:aliasing}
+%\subsection{Aliasing}
+%\label{sec:background:aliasing}
 
-Aliasing is an effect that happens when the camera sensor is unable to correctly represent the patterns and details present in a scene due to its resolution.
+%Aliasing is an effect that happens when the camera sensor is unable to correctly represent the patterns and details present in a scene due to its resolution.
 This often results in artifacts such as Moiré patterns.
-\cite{blog:Wronski2018}.
+\cite{blog:Wronski2018}
 
@@
 
-Generally, super-resolution is a type of technique which increases the resolution of an image.
+Super-resolution is any technique which increases the resolution of an image.
 While there are techniques that work on a single image, we will focus on using multiple frames for super-resolution.
-The main requirements of multi-frame super-resolution are aliased input frames that are sampled at different subpixel offsets (the individual pixel points in each frame captures a different sample of the area it represents)~\cite{Wronski2019}.
+To do this each frame must have some new information to contribute to the final image.
+This is achieved by capturing multiple aliased input frames that are sampled at different subpixel offsets, meaning the individual pixel points in each frame capture a different sample of the area the pixel represents~\cite{Wronski2019}.
 
@@
 
-A kernel (or convolution matrix) in image processing refers to a small matrix that is used to apply effects like blurring or detect features in an image like edges.
+A kernel (or convolution matrix) in image processing refers to a small matrix that is used to apply effects like blurring or to detect features in an image, such as edges.
 The kernel is applied by doing a convolution, adding each pixel of the image to its local neighbors, weighted by the kernel.
@@
 % ----------------------------------------
-\section{Handheld super-resolution}
+\section{Handheld Super-resolution}
 \label{sec:handheldSuperRes}
@@
 
-Wronski et al.~\cite{Wronski2019} introduce an algorithm that uses multiple shifted frames to produce higher resolution images from bursts of underexposed raw frames as part of the smartphone's imaging pipeline.
+Wronski et al.~\cite{Wronski2019} introduced an algorithm that uses multiple shifted frames to produce higher resolution images from bursts of underexposed raw frames as part of the smartphone's imaging pipeline.
 The algorithm is able to directly use Bayer raw frames and removes the need for an explicit demosaicing step in the pipeline.
@@
 This algorithm is used as the merging algorithm in the camera pipeline of the Google Pixel 3 and newer and is what allows for the Pixel's ``super-res zoom" feature.
+This section provides an overview and some results of the algorithm by Wronski et al.~\cite{Wronski2019}.
 
-\subsection{Algorithm overview}
+\subsection{Algorithm Overview}
 
@@
 
-The approach by Wronski et al.~\cite{Wronski2019}, as shown in Figure~\ref{fig:Wronski2019Fig2}, is a process that starts with the the acquisition of a burst from the continuous ring buffer of raw frames in the phone's camera application.
+The approach by Wronski et al., as shown in Figure~\ref{fig:Wronski2019Fig2}, is a process that starts with the the acquisition of a burst from the continuous ring buffer of raw frames in the phone's camera application.
 Next, a single frame is chosen and the rest are aligned to it using a refined version of the algorithm by Hasinoff et al.~\cite{Hasinoff2016}.
@@
 
-\subsection{Hand movement based super-resolution}
+\subsection{Hand Movement Based Super-resolution}
 
@@
 Wronski et al.
-determined this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
+analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography using the rotational measurements from the phone's gyroscope.
+They determined this periodic, random movement while the camera is capturing a burst frames provides sufficient subpixel coverage to create a super-resolution image.
 
-Wronski et al. analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography using the rotational measurements from the phone's gyroscope.
-Their analysis showed that hand movement creates uniformly random angular displacements and relatively slow rotation of the device.
+%Wronski et al. analyzed hand movement in a set of 86 bursts captured by 10 different users during regular smartphone photography using the rotational measurements from the phone's gyroscope.
+%Their analysis showed that hand movement creates uniformly random angular displacements and relatively slow rotation of the device.
 
@@
 
-\subsection{Kernel reconstruction}
-\label{sec:kernelReconstruction}
-
 \begin{figure}
@@
 \end{figure}
+
+\pagebreak
+
+\subsection{Kernel Reconstruction}
+\label{sec:kernelReconstruction}
 
@@
 
-The local pixel weights ($w_{n, i}$) come from a 2D unnormalized anisotropic Gaussian radial basis function kernel which is calculated for the $3\times 3$ neighborhood around the pixel at $(x,y)$.
-It produces a $3\times 3$ matrix of values between 0 and 1 taking into account edges and sharp features in the local area.
+The local pixel weights ($w_{n, i}$) come from a kernel which is calculated for the $3\times 3$ neighborhood around the pixel at $(x,y)$.\footnote{Specifically a 2D unnormalized anisotropic Gaussian radial basis function kernel.} It produces a $3\times 3$ matrix of values between 0 and 1 taking into account edges and sharp features in the local area.
 These weights determine how much each pixel in the $3\times 3$ area will contribute to the target pixel for that frame.
 An example of the matrices that result from the kernel function can be seen in Figure~\ref{fig:Wronski2019Fig8}.
+Here the presence of a sharp feature produces a matrix where more emphasis is put on just the center pixel.
 
@@
 
-\subsection{Motion robustness}
+\pagebreak
+
+\subsection{Motion Robustness}
 \label{sec:robustnessModel}
@@
 It is difficult to reliably align the sequence of images in a burst and even assuming a perfect alignment, changes in the scene and occlusion would still result in some areas of the scene being poorly represented in many frames of the burst.
-This needs to be taken into account to prevent severe artifacting.
+This needs to be taken into account to prevent severe artifacting (Figure~\ref{fig:Wronski2019Fig9} Left).
 To combine frames robustly, a confidence level is assigned to the local neighborhood of each pixel; the map of these confidences is called a \emph{robustness map}.
@@
 
-An example of what this can be seen in the middle image of figure~\ref{fig:Wronski2019Fig9}, the dark areas have a lower robustness value and the lighter areas have a greater robustness value.
+An example of what this can be seen in the middle image of Figure~\ref{fig:Wronski2019Fig9}, where the dark areas have a lower robustness value and the lighter areas have a greater robustness value.
 
@@
 Wronski et al.
-used a variety of methods to evaluate their algorithm and compare it with other demosaicing, merging, and super-resolution techniques.
-They also compared their method as part of a full camera pipeline to that of Hasinoff et al.
-\cite{Hasinoff2016}.
-Figure~\ref{fig:wronskiResults} shows one of these comparison between a traditional method that merges frames of each color channel first and then applies demosaicing to the method developed by Wronski et al.
-\cite{Wronski2019}
+used a variety of methods, both numerical and visual, to evaluate their algorithm and compare it with other demosaicing, merging, and super-resolution techniques.
+They also compared their method as part of a full camera pipeline to that of Hasinoff et al.~\cite{Hasinoff2016}.
+Figure~\ref{fig:wronskiResults} shows one of these comparisons.
 
@@
 The super resolution merge algorithm improves sharpness and detail while eliminating the
-Moiré aliasing artifacts seen in the third (\textbf{red}) frame.~\cite{Wronski2019}}
+Moiré aliasing artifacts (false color banding) seen in the third (\textbf{red}) frame.~\cite{Wronski2019}}
 
@@
 % ----------------------------------------
-\section{Handheld low light photography}
+\section{Handheld Low Light Photography}
 \label{sec:handheldLowLight}
 
-Liba et al.
-\cite{Liba2019} introduce a system for capturing photos in very low light that produces improved color and less noise.
+Liba et al.~\cite{Liba2019} introduce a system for capturing photos in very low light that produces improved color and less noise.
 This system is used for the ``Night Sight" feature on Google Pixel phones.
-The system builds on the existing burst pipeline (from Hasinoff et al.~\cite{Hasinoff2016}) and uses the burst merging algorithm from Wronski et al.~\cite{Wronski2019} on Pixel 3 and newer with modifications to improve low-light photography.
-It uses a \emph{positive-shutter-lag} mode rather than the \emph{zero-shutter-lag} mode used for normal photography to allow for longer exposures.
+The system builds on the existing burst pipeline (from Hasinoff et al.~\cite{Hasinoff2016}) and uses the burst merging algorithm from Wronski et al.~\cite{Wronski2019} with modifications to improve low-light photography.
+It uses a \emph{positive-shutter-lag} mode where the camera waits until the shutter button is pressed to capture the burst of frames.
+This is used rather than the \emph{zero-shutter-lag} mode used for daylight photography to allow for longer exposures.
 
-The main improvements to the camera pipeline are the use of ``motion metering (Section~\ref{sec:motionMetering}) to calculate the settings for each frame based on predicted motion in the scene and camera, a learning-based low-light optimized auto white balance algorithm (Section~\ref{sec:autoWhiteBal}), and tone mapping (Section~\ref{sec:toneMapping}) to produce better colors in low light.
+The main improvements to the camera pipeline from this research are the use of ``motion metering" (Section~\ref{sec:motionMetering}) to calculate the exposure settings for each frame based on predicted motion in the scene and camera, a learning-based low-light optimized auto white balance algorithm (Section~\ref{sec:autoWhiteBal}), and tone mapping (Section~\ref{sec:toneMapping}) to produce better colors in low light.
 
@@
 
-\subsection{Motion metering}
+\subsection{Motion Metering}
 \label{sec:motionMetering}
@@
 It produces a variable exposure schedule that varies based on the amount of motion detected (Figure~\ref{fig:exposure}).
-This is used to select slower exposures for scenes with no motion and shorter exposures for those with motion to reduce motion blur when needed and increase signal-to-noise ratio when possible.
+This is used to select longer exposures for scenes with no motion and shorter exposures for those with motion to reduce motion blur when needed and increase signal-to-noise ratio when possible.
 
@@
 
-\subsection{Auto white balance in low light}
+\subsection{Auto White Balance in Low Light}
 \label{sec:autoWhiteBal}
@@
 
-Humans perceive color correctly of things even when lit with colored light, a phenomenon called color constancy.
-This perception can break down when a photograph is taken under one type of light and viewed under different light.
-It will look tinted (Figure~\ref{fig:whitebalance:default}).
+Humans perceive color correctly even when objects are lit with colored light, an ability called color constancy.
+This perception can break down when a photograph is taken under one type of light and viewed under different light, the image can look tinted (Figure~\ref{fig:whitebalance:default}).
 Cameras correct for this by determining the color of the majority of the illumination in the scene and correcting the colors in the image such that they appear to be lit by a neutral (white) illumination.
-This Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image. \cite{Liba2019, blog:Levoy2018}
+This Automatic White Balance (AWB) step in the camera pipeline is important to produce a pleasing image. \cite{blog:Levoy2018}
 
@@
 %It is a machine learning based algorithm that ``works by learning and then applying a filter to a histogram of log-chrominance pixel and edge intensities"~\cite{Liba2019}. 
-Liba et al. trained the FFCC algorithm with a new dataset and error metric to better handle challenging low-light scenes (Figure~\ref{fig:whitebalance}).
+As described in this section, Liba et al. trained the FFCC algorithm with a new dataset and error metric to better handle challenging low-light scenes (Figure~\ref{fig:whitebalance}).
 
@@
 collected 5000 images using mobile devices of various scenes that demonstrate a range of light levels to be able to train their implementation of FFCC.
-To obtain better ground truth coloring for the training data they had real professional photographers manually white balance the images with the most ``aesthetically preferable" white balance for the scene rather than use a color checker or grey card to empirically measure the ``true" white balance.~\cite{Liba2019}
+To obtain better ground truth coloring for the training data they had real professional photographers manually white balance the images with the most ``aesthetically preferable" white balance for the scene rather than use a color checker or grey card to empirically measure the ``true" white balance.
 
@@
 The issue with traditional error metrics in low light is that they are based on how well the algorithm recovers \emph{all} of the color channels of the illuminant.
-This works well for the close to white true illumination of brighter scenes but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
+This works well for brighter scenes with close to white true illumination but in dark scenes with heavily tinted illuminants the white balanced image may contain pixel values where a single color channel's values are near zero for the whole image.
 Such an image would look the same under all possible transformations of that channel.
-\pagebreak 
-When a color channel is ``missing" like this it can produce inaccurately low accuracy results from the error metric.
-It is also unclear how to set the missing channels in the ground-truth illuminant data.~\cite{Liba2019}
+When a color channel is ``missing" it can produce low accuracy results from the error metric. It is also unclear how to set the missing channels in the ground-truth illuminant data.
 
 The existing error metric for color constancy looks at the error in appearance of a white patch in the image, but that idea doesn't work in heavily-tinted scenes with missing color channels so the improved error metric considers the appearance of an \emph{average} portion of the image under the recovered illumination.
-It is able to be less sensitive to errors in channels with lower mean values in the true image.~\cite{Liba2019}
+It is able to be less sensitive to errors in channels with lower mean values in the true image.
 
@@
 
-\subsection{Tone mapping}
+\pagebreak
+
+\subsection{Tone Mapping}
 \label{sec:toneMapping}
@@
 While human vision loses color sensitivity and spatial acuity as light levels are reduced, Liba et al.
-still wanted to render vibrant and colorful images in low-light rather than emulating human vision.~\cite{Liba2019}
+still wanted to render vibrant and colorful images in low-light rather than emulating human vision.~
 
@@
 For centuries, artists have evoked a nighttime aesthetic through various methods such as the use of darker pigments, suppressed shadows, and increased contrast (Figure~\ref{fig:theOrrery}).
-Liba et al. developed a TMO, inspired by artistic techniques, that maintains vibrant color in dark scenes without looking artificial and while maintaining a nighttime aesthetic.
+Liba et al. developed a TMO, inspired by artistic techniques, that maintains vibrant color in dark scenes without looking artificial while still maintaining a nighttime aesthetic.
 
-Their TMO uses various heuristics on top of the tone mapping of Hasinoff et al.
-\cite{Hasinoff2016}.
-These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a brightness-dependent vignette.
-It produces a more detailed image while maintaining global contrast and the nighttime aesthetic (Figure~\ref{fig:toneMapping:liba}).~\cite{Liba2019}
+Their TMO uses various heuristics on top of the tone mapping of Hasinoff et al.~\cite{Hasinoff2016}.
+These include allowing higher overall gains, limiting the boosting of shadows, allowing compression of higher dynamic ranges, boosting the color saturation inversely to scene brightness, and adding a vignette (darkening around the edges of the image).
+It produces a more detailed image while maintaining global contrast and the nighttime aesthetic (Figure~\ref{fig:toneMapping:liba}).
 
@@
 
-This system was launched in November 2018 on Pixel phones as the ``Night Sight" mode in the camera app.
-On Pixel, Pixel 2, and Pixel 3a it uses a burst merging technique adapted from Hasiniff et al.
+This system by Liba et al.
+was launched in November 2018 on Pixel phones as the ``Night Sight" mode in the camera app.
+On Pixel, Pixel 2, and Pixel 3a it uses a burst merging technique adapted from Hasinoff et al.
 \cite{Hasinoff2016} to better handle motion while on faster Pixel phones it uses the super-resolution merging algorithm described by Wronski et al.
@@
 The comparisons and results here all use the former merging technique.
-\cite{Liba2019}
+
+\pagebreak
 
 Liba et al.
-compared their system to the pipeline it built upon from Hasiniff et al.
+compared their system to the pipeline it built upon from Hasinoff et al.
 \cite{Hasinoff2016}.
-An example of their system producing lower noise, more detail, and more pleasing colors is shown in Figure~\ref{fig:libaResults}.
-They also compared their pipeline with neural network that operates on single raw images by Chen et al.~\cite{Chen2018} using frames from a similar camera to the one the neural network was trained on rather than a smartphone camera and still the system by Liba et al.
-produced higher-quality images and processed images much faster and using less memory.~\cite{Liba2019}
+Figure~\ref{fig:libaResults} shows an example of their system producing lower noise, more detail, and more pleasing colors.
+They also compared their pipeline with a neural network that operates on single raw images by Chen et al.~\cite{Chen2018} using frames from a similar camera to the one the neural network was trained on rather than a smartphone camera and still the system by Liba et al.
+produced higher-quality images and processed images much faster and using less memory.
 
@@
 
-These new computational photography techniques and others like them have allowed Google to continue using the same camera sensor hardware over multiple generations of their Pixel phones while continuing to be highly rated in photo quality.
+Computational photography has allowed for massive advances in smartphone photography, allowing phone cameras to surpass what would be expected from their limited hardware.
+This has allowed many more people than ever before to take high quality photographs with ease.
+The software portion of a camera has become more influential to image quality than the physical hardware.
+These new computational photography techniques have allowed Google to continue using the same camera sensor hardware over multiple generations of their Pixel phones while continuing to be highly rated in photo quality.
 It has also enabled updates that continually improve the camera performance of previous Pixel phones to the point where they compete with the latest from other manufacturers.
-Computational photography techniques will continue to be an ever more important part off cameras and image processing in smartphones, professional photography, and other applications.
+Computational photography techniques will continue to be ever more important in smartphones, professional photography, and other applications.
 
@@
 \label{sec:acknowledgments}
+
+\pagebreak
 
